---
title: "Unpalatability_clean_up"
author: "FIRST AUTHOR INITIALS"
date: "2026-02-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(EnvStats)
library(lme4)
library(gtsummary)
library(emmeans)
library(glmmTMB)
library(prediction)
library(ggeffects)
library(multcomp)
library(multcompView)
library(report)
library(performance)
library(readxl)
library(ggpubr)
library(patchwork)
library(report)
library(parameters)
library(ggnewscale)
library(performance)
library(ggh4x)
source("./plotting_function.R")
```

## Load in the data and meta data for the fishes

```{r, echo=FALSE}
data <- read_xlsx(path = "./2025june11_recopy_data_raw.xlsx",sheet = 1,trim_ws = TRUE,col_names = TRUE)
meta <- read_xlsx(path = "./2025june11_recopy_data_raw.xlsx",sheet = 2,trim_ws = TRUE,col_names = TRUE)
data_merged <- left_join(data,meta,by = "ind_id")

```

## Clean up the data

```{r data clean, echo=FALSE}
data_reduced <- data_merged %>%
  mutate(
    treatment = case_when(treatment == "anesthetized" ~ "anaesthetized", .default = treatment),
    type = case_when(
      type == "luminous" ~ "luminescent",
      type == "nonluminous" ~ "nonluminescent",
      .default = type
    ),
    time_point = factor(time_point,ordered = TRUE,levels = c(1,2,3,4)),
    ind_id = as.factor(ind_id),
    trial_time = as.factor(trial_time),
    food_sample = as.factor(food_sample),
    treatment = as.factor(treatment),
    type = as.factor(type),
    eaten = as.factor(eaten),
    luminescence = as.factor(luminescence),
    species = as.factor(species),
    genus = as.factor(genus),
    ecology = as.factor(ecology),
    days_captive = date - date_captured,
    feeding_id = interaction(ind_id, trial, food_sample, treatment),
    trial_id = interaction(ind_id, trial),
    sample_id = interaction(food_sample, treatment),
    exp_id = interaction(ind_id, trial, order, food_sample, treatment)
  ) %>%
  group_by(ind_id) %>% arrange(trial, order, time_point) %>%
  mutate(
    luminescence_experience = cumsum(replace_na(num_luminescence, 0)),
    photeros_opportunity = case_when(type == "luminescent" ~ 1, .default = NA),
    photeros_experience = cumsum(coalesce(photeros_opportunity, 0)) / 4 + photeros_opportunity *
      0,
    luminescence_experience = case_when(
      luminescence_experience == 0 & photeros_experience > 0 ~ 0,
      .default = luminescence_experience
    ),
    photeros_attack = case_when(
      type == "luminescent" ~ replace_na(num_attacks, 0),
      TRUE ~ 0
      ),
    photeros_attack_experience = cumsum(photeros_attack)
    ) %>%
  ungroup() %>%
  drop_na(eaten) %>%
  mutate(
    num_attacks = case_when(
      eaten == "no" & num_attacks < 1 ~ 0,
      trial_protocol != "pre-trial" &
        eaten == "no" & is.na(num_attacks) ~ 0,
      .default = num_attacks
    ),
    attacked = as.factor(ifelse(num_attacks > 0, "yes", "no")),
    luminescence = case_when(
      is.na(luminescence) &
        type == "luminescent" & treatment == "live" ~ "no",
      .default = luminescence
    ),
    luminescence = case_when(
      type == "luminescent" & treatment == "live" & time_point == 4 ~ NA,
      .default = luminescence
    ),
    rejected = case_when(
      attacked == "yes" & eaten == "no" ~ "Rejected",
      attacked == "yes" &
        eaten == "yes" ~ "Eaten",
      attacked == "no" &
        eaten == "no" ~ "Ignored/Avoided"
    ),
    rejected = as.factor(rejected)
  )  %>%
  arrange(ind_id, trial, order, time_point) %>%
  group_by(trial_protocol, ind_id) %>%
  mutate(min_trial = min(trial), max_trial = max(trial)) %>%
  ungroup() %>% group_by(species, trial) %>%
  mutate(relative_trial = (trial - min_trial) + 1) %>%
  ungroup() %>%
  group_by(exp_id) %>%
  arrange(trial, order, time_point) %>%
  mutate(
    roll_num_attack = cumsum(num_attacks),
    behavior = case_when(
      eaten == "yes" ~ "Eaten",
      eaten == "no" &
        attacked == "yes" ~ "Rejected",
      eaten == "no" &
        attacked == "no" & roll_num_attack == 0 ~ "Unseen",
      eaten == "no" &
        attacked == "no" & roll_num_attack > 0 ~ "Avoided"
    ),
    behavior = as.factor(behavior),
  ) %>% ungroup()

```

## Visualize the attack proportions for each food type, each fish, and the time to consumption for each fish. These are supplementary plots

```{r Supplement Figures}
SFig2 <- data_reduced %>% 
  group_by(trial,food_sample,ind_id) %>%
  summarise(prop_eaten = length(eaten[eaten == "yes"]) / length(eaten)) %>%
  ggplot(aes(x = trial, y = food_sample)) +
  geom_tile(aes(fill = prop_eaten)) + 
  geom_text(aes(label = round(prop_eaten,digits = 2)), color="white",size=2 )+ 
  theme_minimal(base_size = 16) + facet_wrap(ind_id~.) + xlim(1,7)

SFig3 <- data_reduced %>%
  group_by(trial,food_sample,ind_id) %>%
  summarise(prop_attacked = length(attacked[attacked == "yes"]) / length(attacked)) %>%
  ggplot(aes(x = trial, y = food_sample)) +
  geom_tile(aes(fill = prop_attacked)) +
  geom_text(aes(label = round(prop_attacked,digits = 2)), color="white",size=2 )+ 
  theme_minimal(base_size = 16) + facet_wrap(ind_id~.) + xlim(1,7)

plot(SFig2)
plot(SFig3)

data_filtered <-  data_reduced %>% 
  filter(#filtering out the fish that never eat
                        ind_id != "1d",
                        ind_id != "5d",
                        ind_id != "7d",
                        ind_id != "9d") %>%
  filter( # filtering out the trials where the fish don't eat any of the foods
         trial_id != "9a.4",
         trial_id != "3c.5",
         trial_id != "3c.6",
         trial_id != "4c.3")
```

## Making the final data table ##

```{r multi-state data}

msm_data <- data_filtered %>%
  group_by(exp_id) %>%
  arrange(trial,order,time_point) %>%
  mutate(
    roll_num_attack = cumsum(num_attacks),
    behavior = case_when( eaten == "yes" ~ "Eaten",
                          eaten == "no" & attacked == "yes" ~ "Rejected",
                          eaten == "no" & attacked == "no" & roll_num_attack == 0 ~ "Unseen",
                          eaten == "no" & attacked == "no" & roll_num_attack > 0 ~ "Avoided" ),
    behavior = as.factor(behavior),
    behavior = relevel(behavior,ref = "Eaten"),
    behavior_2 = case_when( eaten == "yes" ~ "Eaten",
                          eaten == "no" & attacked == "yes" ~ "Rejected",
                          eaten == "no" & attacked == "no" & roll_num_attack == 0 ~ "Ignored/Unseen"
                          ),
    behavior_2 = as.factor(behavior_2),
    behavior_2 = relevel(behavior_2,ref = "Eaten"),
    behavior_3 = case_when( any(attacked == "yes",na.rm = TRUE) ~ "has attack",.default = "no attack"),
    treatment_2 = case_when(treatment == "frozen" | treatment == "boiled" ~ "dead",.default = treatment),
    treatment_2 = as.factor(treatment_2),
    treatment_2 = relevel(treatment_2, ref = "live"),
    type = relevel(type, ref = "luminescent")
  ) %>% ungroup() %>% filter(type != "control") %>% droplevels()

trial_mean <- mean(msm_data$relative_trial, na.rm = TRUE)
lum_mean   <- mean(msm_data$luminescence_experience, na.rm = TRUE)
order_mean <- mean(msm_data$order, na.rm = TRUE)

msm_data <- msm_data %>%
  mutate(
    relative_trial_c = relative_trial - mean(relative_trial, na.rm = TRUE),
    luminescence_experience_c = luminescence_experience - mean(luminescence_experience, na.rm = TRUE),
    order_c = order - mean(order, na.rm = TRUE),
    order = factor(order,ordered = TRUE,levels = c(1,2,3,4,5,6,7,8,9,10))
  )

msm_data_filtered <- msm_data %>% filter(behavior_3 == "has attack")
```

## Looking at the individual time to event data ##

```{r time to event}

ind_p <- msm_data_filtered %>%
  ggplot(aes(x = time_point, y = ind_id)) +
  geom_tile(aes(fill = eaten)) +
  facet_nested(treatment + type ~ relative_trial) +
  theme_minimal(base_size = 30) +
  coord_cartesian(ratio = 1) +
  theme(panel.grid = element_blank())

ind_p

#ggsave(plot = ind_p,filename = "indidivudal_time_to_eat.jpeg",path = ".",height = 10500,width = 2500,dpi = 300,units = "px")

```

## Looking at the relationship between the occurance of attack and the occurance of a luminous response

```{r Attacks and Luminous Responses}

## how many luminescent events occurred with no attacks?
tb <- msm_data_filtered %>% 
  filter(food_sample == "photeros",treatment == "live",time_point < 4) %>% 
  count(luminescence,attacked) %>%
  pivot_wider(names_from = attacked, values_from = n, values_fill = 0) %>% 
  dplyr::select(-luminescence)

fisher.test(tb) ## p-value = 1.555e-11

tb_long <- msm_data_filtered %>% 
  filter(food_sample == "photeros",treatment == "live",time_point < 4) %>% group_by(genus) %>% 
  count(luminescence,attacked) %>% complete(luminescence,attacked) %>% mutate(n = ifelse(is.na(n),0,n))

library(gridExtra)
library(grid)
library(ggpp)

tb_to_add <- tb_long %>%
  mutate(
    attacked = as.factor(attacked),
    luminescence = as.factor(luminescence),
    N = as.numeric(n),
    luminescence = str_to_title(luminescence),
    attacked = str_to_title(attacked),
    luminescence = fct_relevel(luminescence, c("Yes", "No")),
    attacked = fct_relevel(attacked, c("Yes", "No"))
  ) %>%
  relocate(luminescence, .after = attacked) %>%
  dplyr::select(genus, attacked, luminescence, N) %>%
  arrange(genus, attacked, luminescence) %>%
  rename(
    "Fish attack" = attacked,
    "Luminous response" = luminescence
  )

table_layer_df <- tb_to_add %>%
  group_by(genus) %>%
  tidyr::nest(table_data = c(`Fish attack`, `Luminous response`, N)) %>%
  mutate(
    x = -Inf,
    y = Inf
  )

```

Looking at the relationship between the rate of attack and the rate of luminous response

```{r Looking at attack correlation}

rate_cor_df <- msm_data_filtered %>% 
  group_by(exp_id) %>% 
  mutate("was_eaten" = case_when(any(eaten == "yes") ~ "Eaten",.default = "Rejected")) %>% 
  ungroup() %>% 
  filter(food_sample == "photeros",treatment == "live",time_point < 4,attacked == "yes") %>% 
  group_by(exp_id) %>% arrange(time_point) %>% 
  mutate(num_luminescence = case_when(is.na(num_luminescence) ~ 0,.default = num_luminescence)) %>% droplevels()

rate_model_1 <- glmmTMB(
  num_luminescence ~ num_attacks*genus + time_point + order + (1| ind_id / trial / sample_id ),
  family = poisson(),
  REML = FALSE,
  data = rate_cor_df %>% filter(genus != "Stegastes"),
  na.action = na.fail,
  control = glmmTMBControl(optimizer = optim,optArgs = list(method = "BFGS"))
  )

drop1(rate_model_1,test = "Chisq") ## don't need time_point or order

rate_model_2 <- glmmTMB(
  num_luminescence ~ num_attacks*genus + (1| ind_id / trial / sample_id ),
  family = poisson(),
  REML = FALSE,
  data = rate_cor_df %>% filter(genus != "Stegastes"),
  na.action = na.fail,
  control = glmmTMBControl(optimizer = optim,optArgs = list(method = "BFGS"))
  )

rate_model_3 <- glmmTMB(
  num_luminescence ~ num_attacks*genus + (1| ind_id ),
  family = poisson(),
  REML = FALSE,
  data = rate_cor_df %>% filter(genus != "Stegastes"),
  na.action = na.fail,
  control = glmmTMBControl(optimizer = optim,optArgs = list(method = "BFGS"))
  )

AICcmodavg::aictab(list(rate_model_2,rate_model_3)) ## lower AIC on the 3rd model

test_performance(rate_model_2,rate_model_3) ## more BF support for 3rd model

summary(rate_model_3)
check_model(rate_model_3)
check_overdispersion(rate_model_3)
check_distribution(rate_model_3)
check_zeroinflation(rate_model_3)
plot(effects::allEffects(rate_model_3))
sjPlot::tab_model(rate_model_3,
                  show.intercept = FALSE,
                  show.est = TRUE,
                  show.dev = TRUE,
                  show.loglik = TRUE,
                  show.reflvl = TRUE)

report::report_text(rate_model_3)

library(DHARMa)
sim <- DHARMa::simulateResiduals(rate_model_3)
DHARMa::plotResiduals(rate_model_3)
DHARMa::plotQQunif(rate_model_3)
DHARMa::testZeroInflation(sim)
DHARMa::testUniformity(sim)
DHARMa::testDispersion(sim)
```

Plotting next:

```{r Rate plot}

pred_grid <- rate_cor_df %>% filter(genus != "Stegastes") %>% 
  group_by(genus) %>%
  summarise(
    min_attacks = min(num_attacks, na.rm = TRUE),
    max_attacks = max(num_attacks, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  rowwise() %>%
  mutate(
    num_attacks = list(seq(min_attacks, max_attacks, length.out = 200))
  ) %>%
  unnest(num_attacks) %>%
  mutate(
    genus = factor(genus, levels = levels(rate_cor_df$genus))
  )

pred <- predict(
  rate_model_2,
  newdata = pred_grid,
  type = "link",
  se.fit = TRUE,
  re.form = NA
)

plot_df <- pred_grid %>%
  mutate(
    fit_link = pred$fit,
    se_link = pred$se.fit,
    fit = exp(fit_link),
    lower = exp(fit_link - 1.96 * se_link),
    upper = exp(fit_link + 1.96 * se_link)
  )

my_theme <- ttheme_minimal(
  base_size = 16,
  core = list(
    fg_params = list(fontsize = 14),
    bg_params = list(
      fill = alpha("white"),
      col = NA
    )
  ),
  colhead = list(
    fg_params = list(
      fontsize = 16,
      fontface = "bold"
    ),
    bg_params = list(
      fill = "white",
      col = NA
    )
  )
)

Figure1 <- rate_cor_df %>% filter(was_eaten == "Rejected") %>%
  ggplot(aes(x = num_attacks, y = num_luminescence)) +
  geom_jitter(
    aes(
      shape = was_eaten,
      color = was_eaten,
      fill = was_eaten
    ),
    size = 5,
    stroke = 1.15,
    width = 0.05,height = 0.25
  ) +
  geom_jitter(
    data = rate_cor_df %>% filter(was_eaten == "Eaten"),
    aes(
      shape = was_eaten,
      color = was_eaten,
      fill = was_eaten
    ),
    size = 5,
    stroke = 1.25,
    width = 0.05,height = 0.25
  ) +
  geom_line(
    data = plot_df %>% rename(num_luminescence = fit),
    size = 1,
    color = "#224b5e",
    aes(x = num_attacks, y = num_luminescence)
  ) +
  geom_ribbon(
    data =  plot_df %>% rename(num_luminescence = fit),
    aes(ymin = lower, ymax = upper),
    alpha = 0.25,
    fill = "#224b5e"
  ) +
  theme_bw(base_size = 20) +
  scale_shape_manual(name = "Final result", values = c("Eaten" = 3, "Rejected" = 21)) +
  scale_color_manual(name = "Final result",
                     values = c("Eaten" = "#edc775", "Rejected" = "#94b594")) +
  scale_fill_manual(name = "Final result",
                    values = c("Eaten" = "black", "Rejected" = "white")) +
  xlab("Attack rate\n (fish bites per 20 s)") +
  ylab("Luminescence rate\n (luminous responses per 20 s)") + facet_grid(. ~
                                                                           genus, scales = "free_x") +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "italic")) +
  ggpp::geom_table(
    data = table_layer_df,
    aes(
      x = x,
      y = y,
      label = table_data
    ),table.theme = my_theme,
    inherit.aes = FALSE,
    hjust = -0.025,
    vjust = 1.075,
    size = 1
  )

Figure1

```

Save the figure 1

```{r Save Figure 1}
ggsave(filename = "Figure1.jpeg",plot = Figure1,device = "jpeg",
       path = "/Users/nicholaihensley/Desktop/Burglar Alarm/submissions/biol_let/secondary/figures/",
       width = 19,height = 6.5,units = "in",
       dpi = 300)
```

As a potential suggestion by reviewers and editors, we can perform the same analysis above but removing the outlier @ x = 5.

```{r Removing the high value outlier}

tb_2.1 <- msm_data_filtered %>% 
  filter(food_sample == "photeros",treatment == "live",time_point < 4, num_luminescence < 5 | is.na(num_luminescence)) %>%
  count(luminescence,attacked) %>%
  pivot_wider(names_from = attacked, values_from = n, values_fill = 0) %>% 
  dplyr::select(-luminescence)

fisher.test(tb_2.1) # p-value = 1.74e-11

rate_cor_df_2.1 <- msm_data_filtered %>% 
  group_by(exp_id) %>% 
  mutate("was_eaten" = case_when(any(eaten == "yes") ~ "Eaten",.default = "Rejected")) %>% 
  ungroup() %>% 
  filter(food_sample == "photeros",treatment == "live",time_point < 4,attacked == "yes") %>% 
  mutate(num_luminescence = case_when(is.na(num_luminescence) ~ 0,.default = num_luminescence)) %>% 
  filter(num_luminescence < 5,genus != "Stegastes")

rate_model_2.1 <- glmmTMB(
  num_luminescence ~ num_attacks*genus + (1 | ind_id ),
  family = poisson(),
  REML = FALSE,
  data = rate_cor_df_2.1,
  na.action = na.fail,
  #control = glmmTMBControl(optimizer = nlminb,optArgs = list(method = "Nelder-Mead"))
  )

summary(rate_model_2.1)
check_model(rate_model_2.1)
check_overdispersion(rate_model_2.1)
check_distribution(rate_model_2.1)
check_zeroinflation(rate_model_1)
plot(effects::allEffects(rate_model_2.1))
sjPlot::tab_model(rate_model_2.1,
                  show.intercept = FALSE,
                  show.est = TRUE,
                  show.dev = TRUE,
                  show.loglik = TRUE,
                  show.reflvl = TRUE)

report::report_text(rate_model_2.1)
```

Plotting the supplementary analysis for Figure 1

```{r Plotting the supplementary Fig 1.1}
pred_grid <- rate_cor_df_2.1 %>% #filter(genus != "Stegastes") %>% 
  group_by(genus) %>%
  summarise(
    min_attacks = min(num_attacks, na.rm = TRUE),
    max_attacks = max(num_attacks, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  rowwise() %>%
  mutate(
    num_attacks = list(seq(min_attacks, max_attacks, length.out = 200))
  ) %>%
  unnest(num_attacks) %>%
  mutate(
    genus = factor(genus, levels = levels(rate_cor_df$genus))
  )

pred <- predict(
  rate_model_2.1,
  newdata = pred_grid,
  type = "link",
  se.fit = TRUE,
  re.form = NA
)

plot_df <- pred_grid %>%
  mutate(
    fit_link = pred$fit,
    se_link = pred$se.fit,
    fit = exp(fit_link),
    lower = exp(fit_link - 1.96 * se_link),
    upper = exp(fit_link + 1.96 * se_link)
  )

FigureS2.1 <- rate_cor_df %>% filter(was_eaten == "Rejected",num_luminescence < 5) %>%
  ggplot(aes(x = num_attacks, y = num_luminescence)) +
  geom_jitter(
    aes(
      shape = was_eaten,
      color = was_eaten,
      fill = was_eaten
    ),
    size = 5,
    stroke = 1.15,
    width = 0.05,height = 0.25
  ) +
  geom_jitter(
    data = rate_cor_df %>% filter(was_eaten == "Eaten",num_luminescence < 5),
    aes(
      shape = was_eaten,
      color = was_eaten,
      fill = was_eaten
    ),
    size = 5,
    stroke = 1.25,
    width = 0.05,height = 0.25
  ) +
  geom_line(
    data = plot_df %>% rename(num_luminescence = fit),
    size = 1,
    color = "#224b5e",
    aes(x = num_attacks, y = num_luminescence)
  ) +
  geom_ribbon(
    data =  plot_df %>% rename(num_luminescence = fit),
    aes(ymin = lower, ymax = upper),
    alpha = 0.25,
    fill = "#224b5e"
  ) +
  theme_bw(base_size = 20) +
  scale_shape_manual(name = "Final result", values = c("Eaten" = 3, "Rejected" = 21)) +
  scale_color_manual(name = "Final result",
                     values = c("Eaten" = "#edc775", "Rejected" = "#94b594")) +
  scale_fill_manual(name = "Final result",
                    values = c("Eaten" = "black", "Rejected" = "white")) +
  xlab("Attack rate\n (fish bites per 20 s)") +
  ylab("Luminescence rate\n (luminous responses per 20 s)") + facet_grid(. ~
                                                                           genus, scales = "free_x") +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "italic"))

FigureS2.1

ggsave(filename = "FigureS4_1.jpeg",plot = FigureS2.1,device = "jpeg",
       path = "/Users/nicholaihensley/Desktop/Burglar Alarm/submissions/biol_let/secondary/figures/",
       width = 19,height = 6.5,units = "in",
       dpi = 300)
```

We can also try to perform the same model only including presentations where the food wasn't eaten during the trials (but could still have been eaten in t = 4).

```{r Removing the samples that are eaten during between t = 1,2,3 (n = 9)}
rate_cor_df_2.2 <- rate_cor_df %>% group_by(exp_id) %>% arrange(time_point) %>% 
  mutate(eaten_early = case_when(time_point <= 3 & any(eaten == "yes") ~ "eaten_early",.default = "not")) %>% 
  filter(eaten_early == "not")  %>% 
  filter(num_luminescence < 5,genus!= "Stegastes")

rate_model_2.2 <- glmmTMB(
  num_luminescence ~ num_attacks*genus + (1 | ind_id ),
  family = poisson,
  REML = FALSE,
  data = rate_cor_df_2.2,
  na.action = na.fail,
  control = glmmTMBControl(optimizer = optim,optArgs = list(method = "L-BFGS-B"))
  )

summary(rate_model_2.2)
check_model(rate_model_2.2)
check_overdispersion(rate_model_2.2)
check_distribution(rate_model_2.2)
check_zeroinflation(rate_model_1)
plot(effects::allEffects(rate_model_2.2))
sjPlot::tab_model(rate_model_2.2,
                  show.intercept = FALSE,
                  show.est = TRUE,
                  show.dev = TRUE,
                  show.loglik = TRUE,
                  show.reflvl = TRUE)

report::report_text(rate_model_2.2)
```

Plotting the supplementary analysis for Figure 1.2

```{r Plotting the Supplementary Figure 1.2}
pred_grid <- rate_cor_df_2.2 %>% filter(eaten_early == "not") %>% 
  group_by(genus) %>%
  summarise(
    min_attacks = min(num_attacks, na.rm = TRUE),
    max_attacks = max(num_attacks, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  rowwise() %>%
  mutate(
    num_attacks = list(seq(min_attacks, max_attacks, length.out = 200))
  ) %>%
  unnest(num_attacks) %>%
  mutate(
    genus = factor(genus, levels = levels(rate_cor_df$genus))
  )

pred <- predict(
  rate_model_2.2,
  newdata = pred_grid,
  type = "link",
  se.fit = TRUE,
  re.form = NA
)

plot_df <- pred_grid %>%
  mutate(
    fit_link = pred$fit,
    se_link = pred$se.fit,
    fit = exp(fit_link),
    lower = exp(fit_link - 1.96 * se_link),
    upper = exp(fit_link + 1.96 * se_link)
  )

FigureS2.2 <- rate_cor_df %>% group_by(exp_id) %>% arrange(time_point) %>% 
  mutate(eaten_early = case_when(time_point <= 3 & any(eaten == "yes") ~ "eaten_early",.default = "not")) %>% 
  ungroup() %>% 
  filter(was_eaten == "Rejected",num_luminescence < 5,eaten_early == "not") %>%
  ggplot(aes(x = num_attacks, y = num_luminescence)) +
  geom_jitter(
    aes(
      shape = was_eaten,
      color = was_eaten,
      fill = was_eaten
    ),
    size = 5,
    stroke = 1.15,
    width = 0.05,height = 0.25
  ) +
  geom_jitter(
    data = rate_cor_df %>% filter(was_eaten == "Eaten",num_luminescence < 5),
    aes(
      shape = was_eaten,
      color = was_eaten,
      fill = was_eaten
    ),
    size = 5,
    stroke = 1.25,
    width = 0.05,height = 0.25
  ) +
  geom_line(
    data = plot_df %>% rename(num_luminescence = fit),
    size = 1,
    color = "#224b5e",
    aes(x = num_attacks, y = num_luminescence)
  ) +
  geom_ribbon(
    data =  plot_df %>% rename(num_luminescence = fit),
    aes(ymin = lower, ymax = upper),
    alpha = 0.25,
    fill = "#224b5e"
  ) +
  theme_bw(base_size = 20) +
  scale_shape_manual(name = "Final result", values = c("Eaten" = 3, "Rejected" = 21)) +
  scale_color_manual(name = "Final result",
                     values = c("Eaten" = "#edc775", "Rejected" = "#94b594")) +
  scale_fill_manual(name = "Final result",
                    values = c("Eaten" = "black", "Rejected" = "white")) +
  xlab("Attack rate\n (fish bites per 20 s)") +
  ylab("Luminescence rate\n (luminous responses per 20 s)") + facet_grid(. ~
                                                                           genus, scales = "free_x") +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "italic"))

FigureS2.2

ggsave(filename = "FigureS4_2.jpeg",plot = FigureS2.2,device = "jpeg",
       path = "/Users/nicholaihensley/Desktop/Burglar Alarm/submissions/biol_let/secondary/figures/",
       width = 19,height = 6.5,units = "in",
       dpi = 300)
```

Looking at the correlation between experience and time / trial 

```{r Plotting the correlation between experience and time}
msm_data_filtered %>%
  ggplot(aes(x = time_point, y = ind_id)) +
  geom_point(aes(size = luminescence_experience_c)) +
  facet_nested(relative_trial_c ~ treatment + type) +
  theme_minimal() + facet_grid(relative_trial~genus)
```

The genus Stegastes accumulates very little experience, so I will exclude it from the hazard and attack models downstream

## Looking at the Eaten Model

```{r Eaten Model}

hazard_model_1 <- glmmTMB(
  eaten ~ type + treatment_2 + relative_trial_c + luminescence_experience_c + order + time_point + genus + order:relative_trial_c +
    type:time_point +
    treatment:time_point +
    type:luminescence_experience_c +
    treatment_2:relative_trial_c +
    type:relative_trial_c +
    treatment_2:luminescence_experience_c +
    (1 | ind_id / trial / sample_id ),
  family = binomial("cloglog"),
  REML = FALSE,
  data = msm_data_filtered %>% filter(genus != "Stegastes"), 
  #using the data set where at least one attack occurs in a trial
  na.action = na.fail,
  control = glmmTMBControl(optArgs = list(method = "L-BFGS-B"))
)

drop1(hazard_model_1,test = "Chisq")
```

Drop genus, order, order:trial, type:trial, and treatment:experience are not important without Stegastes in it, so will remove to be consistent with the attack rate model as well, as they don't experience much luminous responses

```{r Eaten Model Round 2}
hazard_model_2 <- glmmTMB(
  eaten ~ type + treatment_2 + relative_trial_c + luminescence_experience_c + time_point +
    # treatment:time_point + # VIF too high to include
    type:luminescence_experience_c +
    treatment_2:relative_trial_c +
    (1 | ind_id / trial / sample_id),
  family = binomial("cloglog"),
  REML = FALSE,
  data = msm_data_filtered %>% filter(genus != "Stegastes"), #using the data set where at least one attack occurs in a trial
  na.action = na.fail,
  control = glmmTMBControl(optArgs = list(method = "L-BFGS-B"),optCtrl = list(iter.max = 500))
)

check_model(hazard_model_2)
check_collinearity(hazard_model_2)
```

Model wouldn't fit without removing the less important random effects. Time point interactions are too collinear to use, so dropping those. 

```{r Eaten Model Comparison}
compare_performance(hazard_model_1,hazard_model_2,rank = TRUE)
plot(compare_performance(hazard_model_1,hazard_model_2,rank = TRUE))
test_performance(hazard_model_1,hazard_model_2)
```

Final model #2 is best for fixed effects. Now testing the random effects best.

```{r Randoem effects comparisons}

hazard_model_2.5 <- glmmTMB(
  eaten ~ type + treatment_2 + relative_trial_c + luminescence_experience_c + time_point + #genus +
    type:luminescence_experience_c +
    treatment_2:relative_trial_c +
    (1 | ind_id / trial / sample_id ),
  family = binomial("cloglog"),
  REML = FALSE,
  data = msm_data_filtered %>% filter(genus != "Stegastes"),
  control = glmmTMBControl(optArgs = list(method = "L-BFGS-B"))
)

hazard_model_3 <- glmmTMB(
  eaten ~ type + treatment_2 + relative_trial_c + luminescence_experience_c + time_point + #genus +
    type:luminescence_experience_c +
    treatment_2:relative_trial_c +
    (1 | ind_id ),
  family = binomial("cloglog"),
  REML = FALSE,
  data = msm_data_filtered %>% filter(genus != "Stegastes"),
  control = glmmTMBControl(optArgs = list(method = "L-BFGS-B"))
)

compare_performance(hazard_model_2.5,hazard_model_3)
plot(compare_performance(hazard_model_2.5,hazard_model_3,rank = TRUE))
test_performance(hazard_model_2.5,hazard_model_3)
```

Hazard model #3 has the best fitting random effects structure

```{r Eaten Model Results}
summary(hazard_model_3)
plot(effects::allEffects(hazard_model_3))
plot(effects::Effect(focal.predictors = "type",mod = hazard_model_3))
plot(effects::Effect(focal.predictors = "treatment_2",mod = hazard_model_3))
plot(effects::Effect(focal.predictors = "time_point",mod = hazard_model_3))
plot(effects::Effect(focal.predictors = "luminescence_experience_c",mod = hazard_model_3))
plot(effects::Effect(focal.predictors = c("relative_trial_c"),mod = hazard_model_3))

car::Anova(hazard_model_3,"II")
check_model(hazard_model_3)
check_collinearity(hazard_model_3)
check_overdispersion(hazard_model_3)
sjPlot::plot_model(hazard_model_3,
                   sort.est = "sort.all",
                   pred.type = "fe",
                   show.values = TRUE,
                   value.offset = -0.35)

sjPlot::tab_model(hazard_model_3,
                  show.intercept = FALSE,
                  show.est = TRUE,
                  show.dev = TRUE,
                  show.loglik = TRUE,
                  show.reflvl = TRUE)

plot(ggpredict(hazard_model_3,
               terms = ~ luminescence_experience_c + relative_trial_c + treatment_2 + type + time_point,
               typical = "marginalmeans",bias_correction = TRUE))

report::report_text(hazard_model_3)

# ## run this part to get the plot of the effects outside R Mardown
# jpeg(filename = "./submissions/biol_let/secondary/figures/Eaten_model_effects.jpeg",
#      width = 20,height = 10,units = "in",res=300)
# plot(effects::allEffects(hazard_model_3),ylab = "Instantaneous risk of being eaten")
# dev.off()
```

Quickly plot the survival function across the time_points for simple visualization

```{r}
pred_grid <- expand.grid(
  type = levels(msm_data_filtered$type),
  treatment_2 = levels(msm_data_filtered$treatment_2),
  relative_trial_c = seq(1:4),
  luminescence_experience_c = 0, #seq(0,max(msm_data_filtered$luminescence_experience)),
  time_point = levels(msm_data_filtered$time_point)
)

pred_grid$time_point <- factor(
  pred_grid$time_point,
  levels = levels(msm_data_filtered$time_point),
  ordered = TRUE
)

X <- model.matrix(
  lme4::nobars(formula(hazard_model_3, component = "cond"))[-2],
  pred_grid
)

beta_hat <- fixef(hazard_model_3)$cond
V_beta <- vcov(hazard_model_3)$cond

set.seed(123)

nsim <- 1000

beta_sim <- MASS::mvrnorm(
  nsim,
  mu = beta_hat,
  Sigma = V_beta
)

eta_sim <- X %*% t(beta_sim)

haz_sim <- 1 - exp(-exp(eta_sim))

sim_df <- bind_cols(
  pred_grid,
  as.data.frame(haz_sim)
)

names(sim_df)[(ncol(pred_grid)+1):ncol(sim_df)] <-
  paste0("sim_", seq_len(nsim))

plot_df <- sim_df %>%
  arrange(type, treatment_2, relative_trial_c, time_point) %>%
  group_by(type, treatment_2, relative_trial_c) %>%
  mutate(
    across(
      starts_with("sim_"),
      ~ 1 - cumprod(1 - .x)
    )
  ) %>%
  ungroup()

plot_df <- plot_df %>%
  rowwise() %>%
  mutate(
    prob = mean(c_across(starts_with("sim_"))),
    lower = quantile(c_across(starts_with("sim_")), 0.025),
    upper = quantile(c_across(starts_with("sim_")), 0.975)
  ) %>%
  ungroup()

plot_df <- plot_df %>%
  mutate(
    condition = interaction(
      str_to_title(type),
      str_to_title(treatment_2)
    )
  )

ggplot(
  plot_df,
  aes(
    x = time_point,
    y = prob,
    colour = condition,
    fill = condition,
    group = condition
  )
) +
  geom_ribbon(
    aes(ymin = lower, ymax = upper),
    alpha = 0.15,
    colour = NA
  ) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  coord_cartesian(ylim = c(0, 1)) +
  labs(
    x = "Time point",
    y = "Cumulative probability of being eaten",
    colour = "Condition",
    fill = "Condition"
  ) +
  theme_bw(base_size = 20) + facet_nested(type + treatment_2 ~ relative_trial_c) + theme(strip.background = element_blank())

```
Will now generate the pairwise comparison table for the supplement, but comparing the relative risks for the different fixed effects

```{r}

## Pairwise comparisons for type and treatments shows in figure 2
emm_type_exp <- emmeans(
  hazard_model_3,
  ~ type + treatment_2 | luminescence_experience_c,
  at = list(
    luminescence_experience_c = 0,
    relative_trial_c = 0
  ),
  type = "response"
)

pairs(emm_type_exp)

## Pairwise slope comparisons for type shown in figure 3

exp_slopes <- emtrends(
  hazard_model_3,
  ~ type,
  var = "luminescence_experience_c"
)

summary(exp_slopes)
pairs(exp_slopes)

## Pairwise slope comparisons for treatment shown in figure 3

trial_slopes <- emtrends(
  hazard_model_3,
  ~ treatment_2,
  var = "relative_trial_c"
)

summary(trial_slopes)
pairs(trial_slopes)
```

Generate the prediciton grid for hazard model and the survival function

```{r}
## ----------------------------
## 1. model-based survival predictions
## ----------------------------

pred_grid <- expand.grid(
  type = levels(msm_data_filtered$type),
  treatment_2 = levels(msm_data_filtered$treatment_2),
  relative_trial_c = 0,
  luminescence_experience_c = 0,
  time_point = levels(msm_data_filtered$time_point)
)

## model matrix for fixed effects
X <- model.matrix(
  delete.response(terms(hazard_model_3)),
  pred_grid
)

## fixed-effect estimates and covariance matrix
beta_hat <- fixef(hazard_model_3)$cond
V_beta <- vcov(hazard_model_3)$cond

## simulate from fixed-effect uncertainty
set.seed(123)

nsim <- 1000

beta_sim <- MASS::mvrnorm(
  n = nsim,
  mu = beta_hat,
  Sigma = V_beta
)

## convert linear predictor -> cloglog hazard
eta_sim <- X %*% t(beta_sim)

haz_sim <- 1 - exp(-exp(eta_sim))

## bind simulations to pred_grid
sim_df <- bind_cols(
  pred_grid,
  as.data.frame(haz_sim)
)

names(sim_df)[(ncol(pred_grid) + 1):ncol(sim_df)] <- paste0("sim_", seq_len(nsim))

## ----------------------------
## 2. hazards -> cumulative probability eaten
## ----------------------------

## ----------------------------
## 2. hazards -> cumulative probability eaten within trial
## ----------------------------

plot_df <- sim_df %>%
  arrange(type, treatment_2, relative_trial_c, time_point) %>%
  group_by(type, treatment_2, relative_trial_c) %>%
  mutate(
    across(
      starts_with("sim_"),
      ~ 1 - cumprod(1 - .x)
    )
  ) %>%
  filter(time_point == levels(time_point)[nlevels(time_point)]) %>%
  ungroup() %>%
  group_by(type, treatment_2) %>%
  summarise(
    across(starts_with("sim_"), mean),
    .groups = "drop"
  ) %>%
  rowwise() %>%
  mutate(
    pred  = mean(c_across(starts_with("sim_"))),
    lower = quantile(c_across(starts_with("sim_")), 0.025),
    upper = quantile(c_across(starts_with("sim_")), 0.975)
  ) %>%
  ungroup() %>%
  mutate(
    type = str_to_title(type),
    treatment_2 = str_to_title(treatment_2),
    type = fct_relevel(type, c("Luminescent", "Nonluminescent")),
    treatment_2 = fct_relevel(treatment_2, c("Live", "Anaesthetized", "Dead")),
    condition = interaction(type, treatment_2)
  )

## ----------------------------
## 3. pairwise contrasts on final cumulative probability eaten
## ----------------------------

sim_long <- plot_df %>%
  dplyr::select(type, treatment_2, condition, starts_with("sim_")) %>%
  tidyr::pivot_longer(
    cols = starts_with("sim_"),
    names_to = "sim",
    values_to = "pred_eaten"
  )

contrast_df <- sim_long %>%
  dplyr::select(condition, sim, pred_eaten) %>%
  tidyr::pivot_wider(
    names_from = condition,
    values_from = pred_eaten
  )

pairwise_contrasts <- combn(levels(plot_df$condition), 2, simplify = FALSE) %>%
  purrr::map_dfr(function(x) {
    diff <- contrast_df[[x[1]]] - contrast_df[[x[2]]]

    tibble::tibble(
      group1 = x[1],
      group2 = x[2],
      estimate = mean(diff),
      lower = quantile(diff, 0.025),
      upper = quantile(diff, 0.975),
      significant = lower > 0 | upper < 0
    )
  })

library(multcompView)

p_mat <- matrix(
  1,
  nrow = nlevels(plot_df$condition),
  ncol = nlevels(plot_df$condition),
  dimnames = list(levels(plot_df$condition), levels(plot_df$condition))
)

for (i in seq_len(nrow(pairwise_contrasts))) {
  p_mat[pairwise_contrasts$group1[i], pairwise_contrasts$group2[i]] <-
    ifelse(pairwise_contrasts$significant[i], 0.01, 1)

  p_mat[pairwise_contrasts$group2[i], pairwise_contrasts$group1[i]] <-
    ifelse(pairwise_contrasts$significant[i], 0.01, 1)
}

letters_df <- multcompView::multcompLetters(
  p_mat,
  threshold = 0.05
)$Letters %>%
  tibble::enframe(
    name = "condition",
    value = ".group"
  )

plot_df <- plot_df %>%
  left_join(letters_df, by = "condition")

## visualizing the distributions of the simulations for each group
sim_long %>% ggplot(aes(x = pred_eaten)) + 
  geom_histogram(aes(color=condition,fill=condition), alpha = 0.15,position = "identity") + 
  theme_bw() + xlab("Cumulative Probability of Being Eaten (n = 1000)") + ylab("Count")
```

Now plotting the bar chart for figure 2

```{r Plotting the bar chat for Figure 2}

## ----------------------------
## 2. raw observed proportions for bars
## ----------------------------

bars_df <- msm_data_filtered %>%
  group_by(exp_id) %>%
  arrange(ind_id, trial, order, time_point, .by_group = TRUE) %>%
  filter(time_point == last(time_point)) %>%
  ungroup() %>%
  mutate(
    sample_id = case_when(
      str_detect(sample_id, "photeros.boiled") ~ "photeros.dead",
      str_detect(sample_id, "photeros.frozen") ~ "photeros.dead",
      str_detect(sample_id, "skogsbergia.boiled") ~ "skogsbergia.dead",
      str_detect(sample_id, "skogsbergia.frozen") ~ "skogsbergia.dead",
      TRUE ~ sample_id
    )
  ) %>%
  group_by(sample_id) %>%
  summarise(
    prop_eaten = mean(eaten == "yes"),
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(
    type = case_when(
      str_detect(sample_id, "photeros") ~ "Luminescent",
      str_detect(sample_id, "skogsbergia") ~ "Nonluminescent",
      TRUE ~ NA_character_
    ),
    treatment_2 = case_when(
      str_detect(sample_id, "live") ~ "Live",
      str_detect(sample_id, "anaesthetized") ~ "Anaesthetized",
      str_detect(sample_id, "dead") ~ "Dead",
      TRUE ~ NA_character_
    ),
    type = fct_relevel(type, c("Luminescent", "Nonluminescent")),
    treatment_2 = fct_relevel(treatment_2, c("Live", "Anaesthetized", "Dead")),
    condition = interaction(type,treatment_2)
  ) %>%
  filter(!is.na(type), !is.na(treatment_2))

## ----------------------------
## 3. plot
## ----------------------------

pd <- position_dodge2(width = 0.75)

Figure2 <- ggplot() +
  geom_col(
    data = bars_df,
    aes(x = treatment_2, y = prop_eaten, fill = condition, colour = condition),
    position = pd,
    alpha = 1,
    width = 0.75,
    linewidth = 0.90
  ) +
  geom_text(
    data = bars_df,
    aes(x = treatment_2, y = -0.05, label = paste0("n = ", n), colour = condition),
    position = pd,
    size = 5
  ) +
  geom_linerange(
    data = plot_df,
    aes(x = treatment_2, ymin = lower, ymax = upper),
    color = "black",
    position = pd,
    linewidth = 1.2
  ) +
  geom_point(
    data = plot_df,
    aes(x = treatment_2, y = pred, fill = condition, shape = condition),
    color = "black",
    position = pd,
    size = 6.5,
    stroke = 1.2
  ) +
  geom_text(
    data = plot_df,
    aes(x = treatment_2, y = 1.05, label = .group),
    position = pd,
    size = 10,
    fontface = "bold"
  ) +

  scale_shape_manual(name = "Food Type",
                     breaks = c('Luminescent.Live',
                                'Luminescent.Anaesthetized',
                                'Luminescent.Dead',
                                'Nonluminescent.Live',
                                'Nonluminescent.Anaesthetized',
                                'Nonluminescent.Dead'
                                ),
                     values = c('Luminescent.Live' = 25,
                                "Nonluminescent.Live" = 24,
                                'Luminescent.Anaesthetized' = 25,
                                'Nonluminescent.Anaesthetized' = 24,
                                'Luminescent.Dead' = 25,
                                'Nonluminescent.Dead' = 24
                                )
  ) +
  
  scale_color_manual(name = "Food Type",
                     breaks = c('Luminescent.Live',
                                'Luminescent.Anaesthetized',
                                'Luminescent.Dead',
                                'Nonluminescent.Live',
                                'Nonluminescent.Anaesthetized',
                                'Nonluminescent.Dead'
                                ),
                     values = c('Luminescent.Live' = "#376795",
                                "Nonluminescent.Live" = "#e76254",
                                'Luminescent.Anaesthetized' = "#528fad",
                                'Nonluminescent.Anaesthetized' = "#ef7a57",
                                'Luminescent.Dead' = "#72bcd5",
                                'Nonluminescent.Dead' = "#f7aa58"
                                )
  ) +
  
  scale_fill_manual(name = "Food Type",
                     breaks = c('Luminescent.Live',
                                'Luminescent.Anaesthetized',
                                'Luminescent.Dead',
                                'Nonluminescent.Live',
                                'Nonluminescent.Anaesthetized',
                                'Nonluminescent.Dead'
                                ),
                     values = c('Luminescent.Live' = "#376795",
                                "Nonluminescent.Live" = "#e76254",
                                'Luminescent.Anaesthetized' = "#528fad",
                                'Nonluminescent.Anaesthetized' = "#ef7a57",
                                'Luminescent.Dead' = "#72bcd5",
                                'Nonluminescent.Dead' = "#f7aa58"
                                )
  ) +
  
  coord_cartesian(ylim = c(-0.07, 1.1)) +
  scale_y_continuous(breaks = c(0.25,0.50,0.75,1)) +
  labs(
    x = "Treatment",
    y = "Proportion eaten (bars)\nCumulative predicted probability of being eaten + 95% CI (triangles + whiskers)"
  ) +
  theme_bw(base_size = 20)

Figure2

```


```{r Saving figure2}

ggsave(plot = Figure2,filename = "Figure2.jpeg", 
       device = "jpeg",
       width = 16,height = 11,units = "in",dpi = 300,
       path = "/Users/nicholaihensley/Desktop/Burglar Alarm/submissions/biol_let/secondary/figures/")

```

Plotting the hazard model other fixed effects

```{r Plotting the Eaten hazard model}
library(MASS)
library(dplyr)
library(tidyr)
library(stringr)
library(forcats)
library(ggplot2)

inv_cloglog <- function(eta) {
  1 - exp(-exp(eta))
}

lum_grid <- seq(
  min(msm_data_filtered$luminescence_experience, na.rm = TRUE),
  max(msm_data_filtered$luminescence_experience, na.rm = TRUE),
  length.out = 200
)

trial_seq <- sort(unique(msm_data_filtered$relative_trial_c))

lum_pred_grid <- expand.grid(
  type = levels(msm_data_filtered$type),
  treatment_2 = levels(msm_data_filtered$treatment_2),
  luminescence_experience_c = lum_grid,
  relative_trial_c = 0,
  time_point = levels(msm_data_filtered$time_point)
)

X <- model.matrix(
  lme4::nobars(formula(hazard_model_3, component = "cond"))[-2],
  lum_pred_grid
)

beta_hat <- fixef(hazard_model_3)$cond
V_beta <- vcov(hazard_model_3)$cond

set.seed(123)
nsim <- 1000

beta_sim <- MASS::mvrnorm(nsim, beta_hat, V_beta)

eta_sim <- X %*% t(beta_sim)
haz_sim <- inv_cloglog(eta_sim)

sim_df <- bind_cols(
  lum_pred_grid,
  as.data.frame(haz_sim)
)

names(sim_df)[(ncol(lum_pred_grid) + 1):ncol(sim_df)] <- paste0("sim_", seq_len(nsim))

lum_df <- sim_df %>%
  arrange(type, treatment_2, luminescence_experience_c, relative_trial_c, time_point) %>%
  group_by(type, treatment_2, luminescence_experience_c, relative_trial_c) %>%
  mutate(
    across(starts_with("sim_"), ~ 1 - cumprod(1 - .x))
  ) %>%
  filter(time_point == levels(time_point)[nlevels(time_point)]) %>%
  ungroup() %>%
  group_by(type, luminescence_experience_c) %>%
  summarise(
    across(starts_with("sim_"), mean),
    .groups = "drop"
  ) %>%
  rowwise() %>%
  mutate(
    prob = mean(c_across(starts_with("sim_"))),
    asymp.LCL = quantile(c_across(starts_with("sim_")), 0.025),
    asymp.UCL = quantile(c_across(starts_with("sim_")), 0.975)
  ) %>%
  ungroup() %>%
  mutate(
    x = luminescence_experience_c,
    type = str_to_title(type)
  )

Fig3.p1 <- ggplot(lum_df, aes(x = x, y = prob, colour = type, fill = type)) +
  geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL), alpha = 0.15, colour = NA) +
  geom_line(linewidth = 1) +
    geom_rug(
    data = subset(msm_data_filtered %>% mutate(type = str_to_title(type)), eaten == "no"),
    aes(x = luminescence_experience, y = 0, color = type),
    position = position_jitter(width = 0.05, height = 0),
    inherit.aes = FALSE,
    sides = "b",
    alpha = 0.25
  ) +
  geom_rug(
    data = subset(msm_data_filtered %>% mutate(type = str_to_title(type)), eaten == "yes"),
    aes(x = luminescence_experience, y = 1, color = type),
    position = position_jitter(width = 0.05, height = 0),
    inherit.aes = FALSE,
    sides = "t",
    alpha = 0.25,
  ) +
  scale_color_manual(name = "Prey Type",
                     values = c("Luminescent" = "#376795",
                                "Nonluminescent" = "#e76254")) +
  scale_fill_manual(name = "Prey Type",
                    values = c("Luminescent" = "#376795",
                               "Nonluminescent" = "#e76254")) +
  labs(
    x = "Luminescence experience",
    y = "Cumulative predicted probability of being eaten",
    colour = "Prey Type",
    fill = "Prey Type"
  ) +
  coord_cartesian(ylim = c(0, 1)) +
  theme_bw(base_size = 20) +
  theme(legend.position = "top")

Fig3.p1
```


```{r}
trial_grid <- seq(
  min(msm_data_filtered$relative_trial, na.rm = TRUE),
  max(msm_data_filtered$relative_trial, na.rm = TRUE),
  length.out = 200
)

trial_pred_grid <- expand.grid(
  type = levels(msm_data_filtered$type),
  treatment_2 = levels(msm_data_filtered$treatment_2),
  relative_trial_c = trial_grid,
  luminescence_experience_c = 0,
  time_point = levels(msm_data_filtered$time_point)
)

X <- model.matrix(
  lme4::nobars(formula(hazard_model_3, component = "cond"))[-2],
  trial_pred_grid
)

eta_sim <- X %*% t(beta_sim)
haz_sim <- inv_cloglog(eta_sim)

sim_df <- bind_cols(
  trial_pred_grid,
  as.data.frame(haz_sim)
)

names(sim_df)[(ncol(trial_pred_grid) + 1):ncol(sim_df)] <- paste0("sim_", seq_len(nsim))

trial_df <- sim_df %>%
  arrange(treatment_2, type, luminescence_experience_c, relative_trial_c, time_point) %>%
  group_by(treatment_2, type, luminescence_experience_c, relative_trial_c) %>%
  mutate(
    across(starts_with("sim_"), ~ 1 - cumprod(1 - .x))
  ) %>%
  filter(
    time_point == levels(time_point)[nlevels(time_point)]
  ) %>% 
  ungroup() %>%
  group_by(treatment_2, relative_trial_c) %>%
  summarise(
    across(starts_with("sim_"), mean),
    .groups = "drop"
  ) %>%
  rowwise() %>%
  mutate(
    prob = mean(c_across(starts_with("sim_"))),
    asymp.LCL = quantile(c_across(starts_with("sim_")), 0.025),
    asymp.UCL = quantile(c_across(starts_with("sim_")), 0.975)
  ) %>%
  ungroup() %>%
  mutate(
    x = relative_trial_c,
    treatment_2 = str_to_title(treatment_2),
    treatment_2 = fct_relevel(treatment_2, c("Live", "Anaesthetized", "Dead"))
  )

Fig3.p2 <- ggplot(trial_df, aes(x = x, y = prob, colour = treatment_2, fill = treatment_2)) +
  geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL), alpha = 0.15, colour = NA) +
  geom_line(linewidth = 1) +
    geom_rug(
    data = subset(msm_data_filtered %>% mutate(treatment_2 = str_to_title(treatment_2)), eaten == "no"),
    aes(x = relative_trial, y = 0, color = treatment_2),
    position = position_jitter(width = 0.15, height = 0),
    inherit.aes = FALSE,
    sides = "b",
    alpha = 0.25
  ) +
  geom_rug(
    data = subset(msm_data_filtered %>% mutate(treatment_2 = str_to_title(treatment_2)), eaten == "yes"),
    aes(x = relative_trial, y = 1, color = treatment_2),
    position = position_jitter(width = 0.15, height = 0),
    inherit.aes = FALSE,
    sides = "t",
    alpha = 0.25,
  ) +
  scale_color_manual(name = "Prey Treatment",
                     values = c("Live" = "#376795",
                                "Anaesthetized" = "#528fad",
                                "Dead" = "#72bcd5")) +
  scale_fill_manual(name = "Prey Treatment",
                    values = c("Live" = "#376795",
                               "Anaesthetized" = "#528fad",
                               "Dead" = "#72bcd5")) +
  labs(
    x = "Trial number",
    y = "Cumulative predicted probability of being eaten",
    colour = "Prey Treatment",
    fill = "Prey Treatment"
  ) +
  coord_cartesian(ylim = c(0, 1)) +
  theme_bw(base_size = 20) +
  theme(legend.position = "top")

Fig3.p2
```

Now plotting Figure 3

```{r}
# Combinding the fig3 plots
Fig3 <- ggarrange(plotlist = c(Fig3.p1,Fig3.p2),ncol = 2,labels = "AUTO",align = "hv",font.label = list(size = 20))

ggsave(plot = Fig3,filename = "Figure3.jpeg", 
       device = "jpeg",
       width = 18,height = 8,units = "in",dpi = 300,
       path = "/Users/nicholaihensley/Desktop/Burglar Alarm/submissions/biol_let/secondary/figures/")
```

## Supplementary Results ##

Plotting all the data even positive controls for the proportion eaten of each type and treatment

```{r Plotting all eaten data for the supplement}

bars_all_df <- data_filtered %>% 
  filter(food_sample != "worm" | food_sample != "cumacean" | food_sample != "mysid") %>% 
  group_by(exp_id) %>%
  arrange(ind_id, trial, order, time_point, .by_group = TRUE) %>%
  filter(time_point == last(time_point)) %>%
  ungroup() %>%
  mutate(
    sample_id = case_when(
      str_detect(sample_id, "photeros.boiled") ~ "photeros.dead",
      str_detect(sample_id, "photeros.frozen") ~ "photeros.dead",
      str_detect(sample_id, "skogsbergia.boiled") ~ "skogsbergia.dead",
      str_detect(sample_id, "skogsbergia.frozen") ~ "skogsbergia.dead",
      str_detect(sample_id, "isopod.control") ~ "isopod.control",
      str_detect(sample_id, "fish.control") ~ "fish.control",
      TRUE ~ sample_id
    )
  ) %>%
  group_by(sample_id) %>%
  summarise(
    prop_eaten = mean(eaten == "yes"),
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(
    type = case_when(
      str_detect(sample_id, "photeros") ~ "Luminescent",
      str_detect(sample_id, "skogsbergia") ~ "Nonluminescent",
      str_detect(sample_id, "isopod") ~ "Isopod",
      str_detect(sample_id, "fish") ~ "Fish",
      TRUE ~ NA_character_
    ),
    treatment_2 = case_when(
      str_detect(sample_id, "live") ~ "Live",
      str_detect(sample_id, "anaesthetized") ~ "Anaesthetized",
      str_detect(sample_id, "dead") ~ "Dead",
      str_detect(sample_id, "control") ~ "Control",
      TRUE ~ NA_character_
    ),
    type = fct_relevel(type, c("Fish","Isopod","Luminescent", "Nonluminescent")),
    treatment_2 = fct_relevel(treatment_2, c("Control","Live", "Anaesthetized", "Dead")),
    condition = interaction(type,treatment_2)
  ) %>%
  filter(!is.na(type), !is.na(treatment_2))

FigureS5 <- ggplot() +
  geom_col(
    data = bars_all_df,
    aes(x = treatment_2, y = prop_eaten, fill = condition, colour = condition),
    position = pd,
    alpha = 1,
    width = 0.75,
    linewidth = 0.90
  ) +
  geom_text(
    data = bars_all_df,
    aes(x = treatment_2, y = -0.05, label = paste0("n = ", n), colour = condition),
    position = pd,
    size = 5
  ) +
  scale_color_manual(name = "Food Type",
                     breaks = c('Fish.Control',
                                'Isopod.Control',
                                'Luminescent.Live',
                                'Luminescent.Anaesthetized',
                                'Luminescent.Dead',
                                'Nonluminescent.Live',
                                'Nonluminescent.Anaesthetized',
                                'Nonluminescent.Dead'
                                ),
                     values = c("Fish.Control" = "grey75",
                                "Isopod.Control" = "grey75",
                                'Luminescent.Live' = "#376795",
                                "Nonluminescent.Live" = "#e76254",
                                'Luminescent.Anaesthetized' = "#528fad",
                                'Nonluminescent.Anaesthetized' = "#ef7a57",
                                'Luminescent.Dead' = "#72bcd5",
                                'Nonluminescent.Dead' = "#f7aa58"
                                )
  ) +
  
  scale_fill_manual(name = "Food Type",
                     breaks = c('Fish.Control',
                                'Isopod.Control',
                                'Luminescent.Live',
                                'Luminescent.Anaesthetized',
                                'Luminescent.Dead',
                                'Nonluminescent.Live',
                                'Nonluminescent.Anaesthetized',
                                'Nonluminescent.Dead'
                                ),
                     values = c("Fish.Control" = "grey75",
                                "Isopod.Control" = "grey75",
                                'Luminescent.Live' = "#376795",
                                "Nonluminescent.Live" = "#e76254",
                                'Luminescent.Anaesthetized' = "#528fad",
                                'Nonluminescent.Anaesthetized' = "#ef7a57",
                                'Luminescent.Dead' = "#72bcd5",
                                'Nonluminescent.Dead' = "#f7aa58"
                                )
  ) +
  
  coord_cartesian(ylim = c(-0.07, 1.1)) +
  scale_y_continuous(breaks = c(0.25,0.50,0.75,1)) +
  labs(
    x = "Treatment",
    y = "Proportion eaten (bars)"
  ) +
  theme_bw(base_size = 20)

FigureS5

ggsave(plot = FigureS5,filename = "FigureS5.jpeg", 
       device = "jpeg",
       width = 13,height = 8,units = "in",dpi = 300,
       path = "/Users/nicholaihensley/Desktop/Burglar Alarm/submissions/biol_let/secondary/figures/")
```

### A separate analysis looking at experience with Photeros during attacks versus luminous signal exposure

```{r Photeros experience model}

msm_photeros <- msm_data_filtered %>% 
    filter(genus != "Stegastes") %>% 
    tidyr::replace_na(list(photeros_attack_experience = 0))

## same as the hazard_model_3 but using the redefined dataset to use the photeros experience without NAs
hazard_model_4 <- glmmTMB(
  eaten ~ type + treatment_2 + relative_trial_c + 
    luminescence_experience_c + 
    # photeros_attack_experience + 
    time_point + 
    type:luminescence_experience_c +
    treatment_2:relative_trial_c +
    (1 | ind_id ),
  family = binomial("cloglog"),
  REML = FALSE,
  data = msm_photeros,
  control = glmmTMBControl(optArgs = list(method = "L-BFGS-B"))
)

check_model(hazard_model_4)
check_collinearity(hazard_model_4)
summary(hazard_model_4)
drop1(hazard_model_4,test="Chisq")

## photeros experience should be dropped if included in the model according to the Chi-sq test

hazard_model_5 <- glmmTMB(
  eaten ~ type + treatment_2 + relative_trial_c + photeros_attack_experience + time_point +
    type:photeros_attack_experience +
    treatment_2:relative_trial_c +
    (1 | ind_id ),
  family = binomial("cloglog"),
  REML = FALSE,
  data = msm_photeros,
  control = glmmTMBControl(optArgs = list(method = "L-BFGS-B"))
)

compare_performance(hazard_model_4,hazard_model_5)
test_performance(hazard_model_4,hazard_model_5)

## in a model with either photeros attack OR luminescence experience
```
The model with the photeros_attack_experience is marginally worse with the BF, but AICc are worse and the R^2 are also worse. This supports that the model with the luminescence_experience is a better predictor of whether food is eaten than simply just attacking / experiencing the food. There does seem to be genus level differences, and we know that from looking at the prey response rates to the two different remaining genera, but the models are equivocal with and without it. Including it does not change the results.


### Looking at the Attack Model now

```{r Attack Model}

attack_model_1 <- glmmTMB(
  attacked ~ type + treatment_2 + relative_trial_c + luminescence_experience_c + order + time_point + genus +
    # type:time_point + #too high of VIF
    # treatment_2:time_point + #too high of VIF
    order:relative_trial_c +
    type:luminescence_experience_c +
    type:relative_trial_c + 
    treatment_2:relative_trial_c +
    treatment_2:luminescence_experience_c +
    (1 | ind_id / trial / sample_id),
  family = binomial("cloglog"),
  REML = FALSE,
  data = msm_data %>% filter(genus != "Stegastes"), #using the entire data set with and without attacks
)

check_collinearity(attack_model_1)
drop1(attack_model_1,test = "Chisq")

# drop1 indicates to get rid of none besides the time_points VIFs interactions


```


```{r Attack Model Results}

summary(attack_model_1)

plot(effects::allEffects(attack_model_1),ylab = "Instantaneous risk of attack")

plot(effects::Effect(focal.predictors = "time_point",
                     mod = attack_model_1),ylab = "Instantaneous risk of attack")
plot(effects::Effect(focal.predictors = c("type","luminescence_experience_c"),
                                          mod = attack_model_1),ylab = "Instantaneous risk of attack")
plot(effects::Effect(focal.predictors = c("type","relative_trial_c"),
                                          mod = attack_model_1),ylab = "Instantaneous risk of attack")
plot(effects::Effect(focal.predictors = c("treatment_2","luminescence_experience_c"),
                     mod = attack_model_1),ylab = "Instantaneous risk of attack")
plot(effects::Effect(focal.predictors = c("treatment_2","relative_trial_c"),
                                          mod = attack_model_1),ylab = "Instantaneous risk of attack")

car::Anova(attack_model_1,"II")
check_model(attack_model_1)
check_collinearity(attack_model_1)
check_overdispersion(attack_model_1)
sjPlot::plot_model(attack_model_1,
                   sort.est = "sort.all",
                   pred.type = "fe",
                   show.values = TRUE,
                   value.offset = -0.35)
```


```{r Attack Model Table}

## to save the effect plot, must run this part outside the markdown
# jpeg(filename = "./submissions/biol_let/secondary/figures/Attack_model_effects.jpeg",
#      width = 20,height = 10,units = "in",res=300)
# plot(effects::allEffects(attack_model_3),ylab = "Instantaneous risk of attack")
# dev.off()

sjPlot::tab_model(attack_model_1,
                  show.intercept = FALSE,
                  show.est = TRUE,
                  show.dev = TRUE,
                  show.loglik = TRUE,
                  show.reflvl = TRUE)

car::Anova(attack_model_1)

report::report(attack_model_1)
```

Plotting the cumulative probability of being eaten for the effects of the attack model as well. I need the following plots:

1) Time point x Order x Genus
2) Type x Experience x Trail 
3) Treatment x Experience x Trial

```{r Attack model: time point}

pred_grid <- expand.grid(
  type = levels(msm_data$type),
  treatment_2 = levels(msm_data$treatment_2),
  relative_trial_c = 0, #seq(1:4),
  luminescence_experience_c = 0, #seq(0,max(msm_data_filtered$luminescence_experience)),
  time_point = levels(msm_data$time_point),
  order = levels(msm_data$order),
  genus = c("Abudefduf","Phaeoptyx")
)

pred_grid$time_point <- factor(
  pred_grid$time_point,
  levels = levels(msm_data$time_point),
  ordered = TRUE
)

X <- model.matrix(
  lme4::nobars(formula(attack_model_1, component = "cond"))[-2],
  pred_grid
)

beta_hat <- fixef(attack_model_1)$cond
V_beta <- vcov(attack_model_1)$cond

set.seed(123)

nsim <- 1000

beta_sim <- MASS::mvrnorm(
  nsim,
  mu = beta_hat,
  Sigma = V_beta
)

eta_sim <- X %*% t(beta_sim)

haz_sim <- 1 - exp(-exp(eta_sim))

sim_df <- bind_cols(
  pred_grid,
  as.data.frame(haz_sim)
)

names(sim_df)[(ncol(pred_grid)+1):ncol(sim_df)] <-
  paste0("sim_", seq_len(nsim))

plot_df <- sim_df %>%
  arrange(genus, type, treatment_2, order, relative_trial_c, luminescence_experience_c, time_point) %>%
  group_by(genus, type, treatment_2, order, relative_trial_c, luminescence_experience_c) %>%
  mutate(
    across(
      starts_with("sim_"),
      ~ 1 - cumprod(1 - .x)
    )
  ) %>%
  ungroup()

plot_df <- plot_df %>%
  rowwise() %>%
  mutate(
    prob = mean(c_across(starts_with("sim_"))),
    lower = quantile(c_across(starts_with("sim_")), 0.025),
    upper = quantile(c_across(starts_with("sim_")), 0.975)
  ) %>%
  ungroup()

plot_df <- plot_df %>%
  mutate(
    condition = interaction(
      str_to_title(type),
      str_to_title(treatment_2)
    )
  )

FigureS6_1 <- ggplot(
  plot_df,
  aes(
    x = time_point,
    y = prob
  )
) +
  geom_ribbon(
    aes(ymin = lower, ymax = upper,fill=condition,group=condition),
    alpha = 0.1
  ) +
  geom_line(linewidth = 1.2,aes(color=condition,group=condition)) +
  geom_point(size = 3,aes(fill=condition),color="black",shape=21) +
  coord_cartesian(ylim = c(0, 1)) +
  labs(
    x = "Time point",
    y = "Cumulative probability of being attacked",
    color = "Type.Treatment",
    fill = "Type.Treatment",
  ) +
  theme_bw(base_size = 20) + facet_grid(genus ~ order) + 
  theme(strip.background = element_blank(),
        strip.text.y = element_text(face = "italic"))

FigureS6_1
```


```{r Attack model: Type x Trial x Experience}

pred_grid <- expand.grid(
  type = levels(msm_data$type),
  treatment_2 = levels(msm_data$treatment_2),
  relative_trial_c = seq(1,4),
  luminescence_experience_c = seq(0,18),
  time_point = levels(msm_data$time_point),
  order = levels(msm_data$order),
  genus = c("Abudefduf","Phaeoptyx")
)

pred_grid$time_point <- factor(
  pred_grid$time_point,
  levels = levels(msm_data$time_point),
  ordered = TRUE
)

X <- model.matrix(
  lme4::nobars(formula(attack_model_1, component = "cond"))[-2],
  pred_grid
)

beta_hat <- fixef(attack_model_1)$cond
V_beta <- vcov(attack_model_1)$cond

set.seed(123)

nsim <- 1000

beta_sim <- MASS::mvrnorm(
  nsim,
  mu = beta_hat,
  Sigma = V_beta
)

eta_sim <- X %*% t(beta_sim)

haz_sim <- 1 - exp(-exp(eta_sim))

sim_df <- bind_cols(
  pred_grid,
  as.data.frame(haz_sim)
)

names(sim_df)[(ncol(pred_grid)+1):ncol(sim_df)] <-
  paste0("sim_", seq_len(nsim))

plot_df <- sim_df %>%
  arrange(genus, type, order, treatment_2, relative_trial_c, luminescence_experience_c, time_point) %>%
  group_by(genus, type, order, treatment_2, relative_trial_c, luminescence_experience_c, ) %>%
  mutate(
    across(starts_with("sim_"), ~ 1 - cumprod(1 - .x))
  ) %>%
  filter(time_point == levels(time_point)[nlevels(time_point)]) %>%
  ungroup() %>%
  group_by(type, luminescence_experience_c, relative_trial_c) %>%
  summarise(
    across(starts_with("sim_"), mean),
    .groups = "drop"
  ) %>%
  rowwise() %>%
  mutate(
    prob  = mean(c_across(starts_with("sim_"))),
    lower = quantile(c_across(starts_with("sim_")), 0.025),
    upper = quantile(c_across(starts_with("sim_")), 0.975)
  ) %>%
  ungroup() %>%
  mutate(
    type = str_to_title(type)
  )

FigureS6_2 <- ggplot(
  plot_df,
  aes(
    x = as.numeric(luminescence_experience_c),
    y = prob,
    colour = type,
    fill = type,
    group = type
  )
) +
  geom_ribbon(aes(ymin = lower, ymax = upper),
              alpha = 0.15,
              colour = NA) +
  geom_line(linewidth = 1.2, aes(color = type)) +
  coord_cartesian(ylim = c(0, 1)) +
  labs(
    x = "Luminescence experience",
    y = "Cumulative probability of being attacked",
    colour = "Type",
    fill = "Type"
  ) +
  theme_bw(base_size = 20) + facet_grid(. ~ as.numeric(relative_trial_c)) + 
  theme(strip.background = element_blank())

FigureS6_2
```


```{r Attack model: Treatment x Trial x Experience}

pred_grid <- expand.grid(
  type = levels(msm_data$type),
  treatment_2 = levels(msm_data$treatment_2),
  relative_trial_c = seq(1,4),
  luminescence_experience_c = seq(0,18),
  time_point = levels(msm_data$time_point),
  order = levels(msm_data$order),
  genus = c("Abudefduf","Phaeoptyx")
)

pred_grid$time_point <- factor(
  pred_grid$time_point,
  levels = levels(msm_data$time_point),
  ordered = TRUE
)

X <- model.matrix(
  lme4::nobars(formula(attack_model_1, component = "cond"))[-2],
  pred_grid
)

beta_hat <- fixef(attack_model_1)$cond
V_beta <- vcov(attack_model_1)$cond

set.seed(123)

nsim <- 1000

beta_sim <- MASS::mvrnorm(
  nsim,
  mu = beta_hat,
  Sigma = V_beta
)

eta_sim <- X %*% t(beta_sim)

haz_sim <- 1 - exp(-exp(eta_sim))

sim_df <- bind_cols(
  pred_grid,
  as.data.frame(haz_sim)
)

names(sim_df)[(ncol(pred_grid)+1):ncol(sim_df)] <-
  paste0("sim_", seq_len(nsim))

plot_df <- sim_df %>%
  arrange(genus, type, order, treatment_2, relative_trial_c, luminescence_experience_c, time_point) %>%
  group_by(genus, type, order, treatment_2, relative_trial_c, luminescence_experience_c, ) %>%
  mutate(
    across(starts_with("sim_"), ~ 1 - cumprod(1 - .x))
  ) %>%
  filter(time_point == levels(time_point)[nlevels(time_point)]) %>%
  ungroup() %>%
  group_by(treatment_2, luminescence_experience_c, relative_trial_c) %>%
  summarise(
    across(starts_with("sim_"), mean),
    .groups = "drop"
  ) %>%
  rowwise() %>%
  mutate(
    prob  = mean(c_across(starts_with("sim_"))),
    lower = quantile(c_across(starts_with("sim_")), 0.025),
    upper = quantile(c_across(starts_with("sim_")), 0.975)
  ) %>%
  ungroup() %>%
  mutate(
    type = str_to_title(treatment_2)
  )

FigureS6_3 <- ggplot(
  plot_df,
  aes(
    x = as.numeric(luminescence_experience_c),
    y = prob,
    colour = treatment_2,
    fill = treatment_2,
    group = treatment_2
  )
) +
  geom_ribbon(aes(ymin = lower, ymax = upper),
              alpha = 0.15,
              colour = NA) +
  geom_line(linewidth = 1.2, aes(color = treatment_2)) +
  coord_cartesian(ylim = c(0, 1)) +
  labs(
    x = "Luminescence experience",
    y = "Cumulative probability of being attacked",
    colour = "Treatment",
    fill = "Treatment"
  ) +
  theme_bw(base_size = 20) + facet_grid(. ~ as.numeric(relative_trial_c)) + 
  theme(strip.background = element_blank())

FigureS6_3
```

```{r Saving them as a plot}
Fig6_row1 <- ggarrange(FigureS6_1,ncol = 1,nrow = 1,legend = "top")
Fig6_row2 <- ggarrange(FigureS6_2,FigureS6_3,nrow = 2,ncol = 1,legend = "top")
FigS6 <- ggarrange(Fig6_row1,Fig6_row2,ncol = 1)
FigS6
ggsave(filename = "FigureS6.jpeg",plot = FigS6,device = "jpeg",
       path = "./submissions/biol_let/secondary/figures/",width = 15,height = 18,units = "in",dpi = 300)
```




