---
title: "Unpalatability_clean_up"
author: "FIRST AUTHOR INITIALS"
date: "2026-02-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(EnvStats)
library(lme4)
library(gtsummary)
library(emmeans)
library(glmmTMB)
library(prediction)
library(ggeffects)
library(multcomp)
library(multcompView)
library(report)
library(performance)
library(readxl)
library(ggpubr)
library(patchwork)
library(report)
library(parameters)
library(ggnewscale)
library(performance)
library(ggh4x)
```

## Load in the data and meta data for the fishes

```{r, echo=FALSE}

data <- read_xlsx(
  path = "./2025june11_recopy_data_raw.xlsx",
  sheet = 1,
  trim_ws = TRUE,
  col_names = TRUE
)
meta <- read_xlsx(
  path = "./2025june11_recopy_data_raw.xlsx",
  sheet = 2,
  trim_ws = TRUE,
  col_names = TRUE
)
data_merged <- left_join(data, meta, by = "ind_id")

```

## Clean up the data

```{r data clean, echo=FALSE}


data_reduced <- data_merged %>%
  mutate(
    treatment = case_when(treatment == "anesthetized" ~ "anaesthetized", .default = treatment),
    type = case_when(
      type == "luminous" ~ "luminescent",
      type == "nonluminous" ~ "nonluminescent",
      .default = type
    ),
    time_point = factor(time_point, ordered = TRUE, levels = c(1, 2, 3, 4)),
    ind_id = as.factor(ind_id),
    trial_time = as.factor(trial_time),
    food_sample = as.factor(food_sample),
    treatment = as.factor(treatment),
    type = as.factor(type),
    eaten = as.factor(eaten),
    luminescence = as.factor(luminescence),
    species = as.factor(species),
    genus = as.factor(genus),
    ecology = as.factor(ecology),
    days_captive = date - date_captured,
    feeding_id = interaction(ind_id, trial, food_sample, treatment),
    trial_id = interaction(ind_id, trial),
    sample_id = interaction(food_sample, treatment),
    exp_id = interaction(ind_id, trial, order, food_sample, treatment)
  ) %>%
  group_by(ind_id) %>% arrange(trial, order, time_point) %>%
  mutate(
    luminescence_experience = cumsum(replace_na(num_luminescence, 0)),
    photeros_opportunity = case_when(type == "luminescent" ~ 1, .default = NA),
    photeros_experience = cumsum(coalesce(photeros_opportunity, 0)) / 4 + photeros_opportunity *
      0,
    luminescence_experience = case_when(
      luminescence_experience == 0 & photeros_experience > 0 ~ 0,
      .default = luminescence_experience
    ),
    photeros_attack = case_when(type == "luminescent" ~ replace_na(num_attacks, 0), TRUE ~ 0),
    photeros_attack_experience = cumsum(photeros_attack)
  ) %>%
  ungroup() %>%
  drop_na(eaten) %>%
  mutate(
    num_attacks = case_when(
      eaten == "no" & num_attacks < 1 ~ 0,
      trial_protocol != "pre-trial" &
        eaten == "no" & is.na(num_attacks) ~ 0,
      .default = num_attacks
    ),
    attacked = as.factor(ifelse(num_attacks > 0, "yes", "no")),
    luminescence = case_when(
      is.na(luminescence) &
        type == "luminescent" & treatment == "live" ~ "no",
      .default = luminescence
    ),
    luminescence = case_when(
      type == "luminescent" & treatment == "live" & time_point == 4 ~ NA,
      .default = luminescence
    ),
    rejected = case_when(
      attacked == "yes" & eaten == "no" ~ "Rejected",
      attacked == "yes" &
        eaten == "yes" ~ "Eaten",
      attacked == "no" &
        eaten == "no" ~ "Ignored/Avoided"
    ),
    rejected = as.factor(rejected)
  )  %>%
  arrange(ind_id, trial, order, time_point) %>%
  group_by(trial_protocol, ind_id) %>%
  mutate(min_trial = min(trial), max_trial = max(trial)) %>%
  ungroup() %>% group_by(species, trial) %>%
  mutate(relative_trial = (trial - min_trial) + 1) %>%
  ungroup() %>%
  group_by(exp_id) %>%
  arrange(trial, order, time_point) %>%
  mutate(
    roll_num_attack = cumsum(num_attacks),
    behavior = case_when(
      eaten == "yes" ~ "Eaten",
      eaten == "no" &
        attacked == "yes" ~ "Rejected",
      eaten == "no" &
        attacked == "no" & roll_num_attack == 0 ~ "Unseen",
      eaten == "no" &
        attacked == "no" & roll_num_attack > 0 ~ "Avoided"
    ),
    behavior = as.factor(behavior),
  ) %>% ungroup()

```

## Visualize the attack proportions for each food type, each fish, and the time to consumption for each fish. These are supplementary plots

```{r Supplemental Figures that show the data and set -up}


### read in the images for the set-up, and generate the single plot
library(magick)
si5.1 <- image_read("submissions/plots/supplemental_figure1.1.jpg")
sfig5.1 <- image_ggplot(image_trim(si5.1))

si5.2 <- image_read("submissions/plots/supplemental_figure1.2.jpg")
sfig5.2 <- image_ggplot(image_trim(si5.2))
SFig5 <- ggarrange(
  sfig5.1,
  sfig5.2,
  ncol = 1,
  labels = "AUTO",
  align = "h",
  font.label = list(size = 18)
)

## look at the pre-trials with just the other foods
SFig4 <- data_reduced %>% 
  filter(
    trial_protocol == "water protocol",
    food_sample != "photeros",
    food_sample != "skogsbergia",
    genus != "Stegastes"
  ) %>%
  group_by(trial, ind_id) %>% filter(any(eaten == "yes")) %>%
  ungroup() %>%
  group_by(genus, food_sample, eaten) %>%
  tally %>%
  spread(eaten, n, fill = 0) %>% ggplot(aes(x = food_sample, y = yes / (yes + no))) +
  geom_col(aes(col = food_sample), fill = "white") +
  geom_label(aes(label = paste0("n = ", yes + no), col = food_sample)) +
  theme_bw(base_size = 20) + xlab("Potential positive control foods") + 
  ylab("Proportion of preliminary tests when eaten") +
  theme(legend.position = "none") + facet_grid(. ~ genus) + 
  theme(strip.text.x = element_text(face = "italic")) + ylim(c(0,1))

## visualize the proportion of presentations per trial where food was eaten for each individual

SFig2 <- data_reduced %>% filter(genus != "Stegastes",trial_protocol == "ostracod protocol") %>%
  group_by(genus, fish_id, relative_trial, exp_id) %>%
  summarise(eaten_item = as.numeric(any(eaten == "yes", na.rm = TRUE)), .groups = "drop") %>%
  group_by(genus, fish_id, relative_trial) %>%
  summarise(
    prop_eaten = mean(eaten_item),
    n_food_items = n(),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = relative_trial, y = fish_id)) +
  geom_tile(aes(fill = prop_eaten),color="black") +
  geom_text(aes(label = round(prop_eaten, digits = 2)), color = "white", size = 5) +
  theme_bw(base_size = 16) + facet_wrap(genus ~ .) + #coord_cartesian(xlim = c(1, 4)) +
  ylab("Individual Fish ID") +
  xlab("Trial number") + labs(subtitle = "Proportion of prey eaten within a trial") +
  scale_fill_continuous(name = "Proportion eaten") + 
  theme(strip.text.x = element_text(face = "italic"))

SFig3 <- data_reduced %>% filter(genus != "Stegastes", trial_protocol == "ostracod protocol") %>%
  group_by(genus, fish_id, relative_trial, exp_id) %>%
  summarise(attacked_item = as.numeric(any(attacked == "yes", na.rm = TRUE)), .groups = "drop") %>%
  group_by(genus, fish_id, relative_trial) %>%
  summarise(
    prop_attacked = mean(attacked_item),
    n_food_items = n(),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = relative_trial, y = fish_id)) +
  geom_tile(aes(fill = prop_attacked),color="black") +
  geom_text(aes(label = round(prop_attacked, digits = 2)), color = "white", size =
              5) +
  theme_bw(base_size = 16) + facet_wrap(genus ~ .) + #coord_cartesian(xlim = c(1, 4)) +
  ylab("Individual Fish ID") +
  xlab("Trial number") + 
  labs(subtitle = "Proportion of prey attacked within a trial") +
  scale_fill_continuous(name = "Proportion attacked") + theme(strip.text.x = element_text(face = "italic"))

ggsave(
  SFig5,
  filename = "./submissions/plots/Sfig5.jpeg",
  device = "jpeg",
  dpi = 300,
  units = "in",
  height = 14,
  width = 12,
  limitsize = TRUE
)

ggsave(
  plot = SFig4,
  filename = "./submissions/plots/Sfig4.jpeg",
  device = "jpeg",
  dpi = 300,
  units = "in",
  width = 18,
  height = 8
)


SFig3tot <- ggarrange(SFig3,SFig2,ncol = 1,nrow = 2,align = "h",font.label = list(size=20),labels = "AUTO")


ggsave(
  plot = SFig3tot,
  filename = "./submissions/plots/Sfig3tot.jpeg",
  device = "jpeg",
  dpi = 300,
  units = "in",
  width = 16,
  height = 16
)


```

## Making the final data table

```{r multi-state data}

data_filtered <- data_reduced %>%
  group_by(exp_id) %>%
  arrange(time_point) %>%
  filter(any(attacked == "yes"))

msm_data <- data_filtered %>%
  group_by(exp_id) %>%
  arrange(trial, order, time_point) %>%
  mutate(
    roll_num_attack = cumsum(num_attacks),
    behavior = case_when(
      eaten == "yes" ~ "Eaten",
      eaten == "no" &
        attacked == "yes" ~ "Rejected",
      eaten == "no" &
        attacked == "no" & roll_num_attack == 0 ~ "Unseen",
      eaten == "no" &
        attacked == "no" & roll_num_attack > 0 ~ "Avoided"
    ),
    behavior = as.factor(behavior),
    behavior = relevel(behavior, ref = "Eaten"),
    behavior_2 = case_when(
      eaten == "yes" ~ "Eaten",
      eaten == "no" &
        attacked == "yes" ~ "Rejected",
      eaten == "no" &
        attacked == "no" & roll_num_attack == 0 ~ "Ignored/Unseen"
    ),
    behavior_2 = as.factor(behavior_2),
    behavior_2 = relevel(behavior_2, ref = "Eaten"),
    behavior_1.3 = case_when(any(attacked == "yes", na.rm = TRUE) ~ "has attack", .default = "no attack"),
    treatment_2 = case_when(
      treatment == "frozen" ~ "dead",
      treatment == "boiled" ~ "dead",
      .default = treatment
    ),
    treatment_2 = as.factor(treatment_2),
    treatment_2 = relevel(treatment_2, ref = "live"),
    type = relevel(type, ref = "luminescent")
  ) %>% ungroup() %>% filter(type != "control") %>% droplevels()

trial_mean <- mean(msm_data$relative_trial, na.rm = TRUE)
lum_mean   <- mean(msm_data$luminescence_experience, na.rm = TRUE)
order_mean <- mean(msm_data$order, na.rm = TRUE)

msm_data <- msm_data %>%
  mutate(
    relative_trial_c = relative_trial - mean(relative_trial, na.rm = TRUE),
    luminescence_experience_c = luminescence_experience - mean(luminescence_experience, na.rm = TRUE),
    order_c = order - mean(order, na.rm = TRUE),
    order = factor(
      order,
      ordered = TRUE,
      levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
    )
  )

msm_data_filtered <- msm_data %>% filter(behavior_1.3 == "has attack")

```

## Looking at the individual time to event data, which is the raw data for the analyses

```{r time to event}


ind_p_att <- msm_data_filtered %>% filter(genus != "Stegastes") %>% 
  mutate(type = case_when(type == "luminescent" ~ "bioluminescent",.default = type),
    treatment = factor(
    treatment,
    labels = c("live", "anaesthetized", "frozen", "boiled")
  )) %>%
  ggplot(aes(x = time_point, y = fish_id)) +
  geom_tile(aes(fill = attacked), color = "grey") +
  facet_nested(
    treatment + type ~ genus + relative_trial,
    nest_line = element_line(),
    strip = strip_nested(
      by_layer_x = TRUE,
      text_x = list(element_text(face = "italic"),NULL) # Changed NA to NULL
    )
  ) +
  theme_bw(base_size = 18) +
  theme(panel.grid = element_blank(),
        strip.background = element_blank()
        ) +
  ylab("Individual Fish ID") +
  xlab("Time Point") +
  labs(title = "Time to event (prey attack) for each fish across trials",
       subtitle = "Separated by prey type and treatment") +
  scale_fill_manual(name = "\u2265 1 Attack Observed",
                    values = c("yes" = "black", "no" = "grey"))

ggsave(
  plot = ind_p_att,
  filename = "indidivudal_time_to_attack.jpeg",
  path = "./submissions/plots/",
  height = 5000,
  width = 5000,
  dpi = 300,
  units = "px"
)

ind_p_eat <- msm_data_filtered %>% filter(genus != "Stegastes") %>% 
  mutate(type = case_when(type == "luminescent" ~ "bioluminescent",.default = type),
    treatment = factor(
    treatment,
    labels = c("live", "anaesthetized", "frozen", "boiled")
  )) %>%
  ggplot(aes(x = time_point, y = fish_id)) +
  geom_tile(aes(fill = eaten), color = "grey") +
  facet_nested(
    treatment + type ~ genus + relative_trial,
    nest_line = element_line(),
    strip = strip_nested(
      by_layer_x = TRUE,
      text_x = list(element_text(face = "italic"),NULL) # Changed NA to NULL
    )
  ) +
  theme_bw(base_size = 18) +
  theme(panel.grid = element_blank(),
        strip.background = element_blank()
        ) +
  ylab("Individual Fish ID") +
  xlab("Time Point") +
  labs(title = "Time to event (prey eaten) for each fish across trials",
       subtitle = "Separated by prey type and treatment") +
  scale_fill_manual(name = "Prey Eaten",
                    values = c("yes" = "black", "no" = "grey"))

ggsave(
  plot = ind_p_eat,
  filename = "indidivudal_time_to_eat.jpeg",
  path = "./submissions/plots/",
  height = 5000,
  width = 5000,
  dpi = 300,
  units = "px"
)

```

## Looking at the relationship between the occurrence of attack and the occurrence of a luminous response

```{r Attacks and Luminous Responses}


## how many luminescent events occurred with no attacks?
tb <- msm_data_filtered %>%
  filter(food_sample == "photeros",
         treatment == "live",
         time_point < 4,
         genus != "Stegastes") %>%
  count(luminescence, attacked) %>%
  pivot_wider(names_from = attacked,
              values_from = n,
              values_fill = 0) %>%
  dplyr::select(-luminescence)

fisher.test(tb) ## p-value = 3.139e-12

tb_long <- msm_data_filtered %>%
  filter(food_sample == "photeros",
         treatment == "live",
         time_point < 4,
         genus != "Stegastes") %>%
  group_by(genus) %>%
  count(luminescence, attacked) %>% complete(luminescence, attacked) %>% mutate(n = ifelse(is.na(n), 0, n))

library(gridExtra)
library(grid)
library(ggpp)

tb_to_add <- tb_long %>%
  mutate(
    attacked = as.factor(attacked),
    luminescence = as.factor(luminescence),
    N = as.numeric(n),
    luminescence = str_to_title(luminescence),
    attacked = str_to_title(attacked),
    luminescence = fct_relevel(luminescence, c("Yes", "No")),
    attacked = fct_relevel(attacked, c("Yes", "No"))
  ) %>%
  relocate(luminescence, .after = attacked) %>%
  dplyr::select(genus, attacked, luminescence, N) %>%
  arrange(genus, attacked, luminescence) %>%
  rename("Fish\nattack" = attacked, "Bioluminescent\nresponse" = luminescence)

table_layer_df <- tb_to_add %>%
  group_by(genus) %>%
  tidyr::nest(table_data = c(`Fish\nattack`, `Bioluminescent\nresponse`, N)) %>%
  mutate(x = -Inf, y = Inf)

```

Looking at the relationship between the rate of attack and the rate of luminous response

```{r Looking at attack correlation}


rate_cor_df <- msm_data_filtered %>% filter(genus != "Stegastes") %>%
  group_by(exp_id) %>%
  mutate("was_eaten" = case_when(any(eaten == "yes") ~ "Eaten", .default = "Rejected")) %>%
  ungroup() %>%
  filter(food_sample == "photeros",
         treatment == "live",
         time_point < 4,
         attacked == "yes") %>%
  group_by(exp_id) %>% arrange(time_point) %>%
  mutate(num_luminescence = case_when(is.na(num_luminescence) ~ 0, .default = num_luminescence)) %>% droplevels()

rate_model_1 <- glmmTMB(
  num_luminescence ~ num_attacks * genus + time_point + order + (1 |
                                                                   ind_id / trial / sample_id),
  family = poisson(),
  REML = FALSE,
  data = rate_cor_df %>% filter(genus != "Stegastes"),
  na.action = na.fail,
  control = glmmTMBControl(optimizer = optim, optArgs = list(method = "BFGS"))
)

drop1(rate_model_1, test = "Chisq") ## don't need time_point or order

rate_model_2 <- glmmTMB(
  num_luminescence ~ num_attacks * genus + (1 |
                                              ind_id / trial / sample_id),
  family = poisson(),
  REML = FALSE,
  data = rate_cor_df %>% filter(genus != "Stegastes"),
  na.action = na.fail,
  control = glmmTMBControl(optimizer = optim, optArgs = list(method = "BFGS"))
)

rate_model_1.3 <- glmmTMB(
  num_luminescence ~ num_attacks * genus + (1 | ind_id),
  family = poisson(),
  REML = FALSE,
  data = rate_cor_df %>% filter(genus != "Stegastes"),
  na.action = na.fail,
  control = glmmTMBControl(optimizer = optim, optArgs = list(method = "BFGS"))
)

AICcmodavg::aictab(list(rate_model_2, rate_model_1.3)) ## lower AIC on the 3rd model

test_performance(rate_model_2, rate_model_1.3) ## more BF support for 3rd model

summary(rate_model_1.3)
check_model(rate_model_1.3)
check_overdispersion(rate_model_1.3)
check_distribution(rate_model_1.3)
check_zeroinflation(rate_model_1.3)
plot(effects::allEffects(rate_model_1.3))
sjPlot::tab_model(
  rate_model_1.3,
  show.intercept = FALSE,
  show.est = TRUE,
  show.dev = TRUE,
  show.loglik = TRUE,
  show.reflvl = TRUE
)

report::report_text(rate_model_1.3)

library(DHARMa)
sim <- DHARMa::simulateResiduals(rate_model_1.3)
DHARMa::plotResiduals(rate_model_1.3)
DHARMa::plotQQunif(rate_model_1.3)
DHARMa::testZeroInflation(sim)
DHARMa::testUniformity(sim)
DHARMa::testDispersion(sim)

```

Plotting next:

```{r Rate plot}


pred_grid <- rate_cor_df %>% filter(genus != "Stegastes") %>%
  group_by(genus) %>%
  summarise(
    min_attacks = min(num_attacks, na.rm = TRUE),
    max_attacks = max(num_attacks, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  rowwise() %>%
  mutate(num_attacks = list(seq(min_attacks, max_attacks, length.out = 200))) %>%
  unnest(num_attacks) %>%
  mutate(genus = factor(genus, levels = levels(rate_cor_df$genus)))

pred <- predict(
  rate_model_1.3,
  newdata = pred_grid,
  type = "link",
  se.fit = TRUE,
  re.form = NA
)

plot_df <- pred_grid %>%
  mutate(
    fit_link = pred$fit,
    se_link = pred$se.fit,
    fit = exp(fit_link),
    lower = exp(fit_link - 1.96 * se_link),
    upper = exp(fit_link + 1.96 * se_link)
  )

my_theme <- ttheme_minimal(
  base_size = 16,
  core = list(
    fg_params = list(fontsize = 14),
    bg_params = list(fill = alpha("white"), col = NA)
  ),
  colhead = list(
    fg_params = list(fontsize = 16, fontface = "bold"),
    bg_params = list(fill = "white", col = NA)
  )
)

Figure1 <- rate_cor_df %>% filter(#was_eaten == "Rejected",
                                  genus != "Stegastes") %>%
  ggplot(aes(x = num_attacks, y = num_luminescence)) +
  geom_jitter(
    aes(
      shape = was_eaten,
      color = was_eaten,
      fill = was_eaten
    ),
    size = 5,
    stroke = 1.15,
    width = 0.05,
    height = 0.25
  ) +
  geom_jitter(
    data = rate_cor_df %>% filter(was_eaten == "Eaten"),
    aes(
      shape = was_eaten,
      color = was_eaten,
      fill = was_eaten
    ),
    size = 5,
    stroke = 1.25,
    width = 0.05,
    height = 0.25
  ) +
  geom_line(
    data = plot_df %>% rename(num_luminescence = fit),
    size = 1,
    color = "#376795",
    aes(x = num_attacks, y = num_luminescence)
  ) +
  geom_ribbon(
    data =  plot_df %>% rename(num_luminescence = fit),
    aes(ymin = lower, ymax = upper),
    alpha = 0.25,
    fill = "#376795"
  ) +
  theme_bw(base_size = 20) +
  scale_shape_manual(name = "Final result", values = c("Eaten" = 3, "Rejected" = 21)) +
  scale_color_manual(name = "Final result",
                     values = c("Eaten" = "black", "Rejected" = "grey")) +
  scale_fill_manual(name = "Final result",
                    values = c("Eaten" = "black", "Rejected" = "white")) +
  xlab("Attack rate\n (fish bites per 20 s)") +
  ylab("Bioluminescence rate\n (responses per 20 s)") + facet_grid(. ~ genus) +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "italic")) +
  ggpp::geom_table(
    data = table_layer_df,
    aes(x = x, y = y, label = table_data),
    table.theme = my_theme,
    inherit.aes = FALSE,
    hjust = -0.025,
    vjust = 1.075,
    size = 1
  )

## pull in the images
library(magick)
i1 <- image_read("submissions/plots/frame_0922.jpg")
plot(i1)
gi1 <- image_ggplot(image_trim(i1))

i2 <- image_read("submissions/plots/frame_0923.jpg")
plot(i2)
gi2 <- image_ggplot(image_trim(i2))

i3 <- image_read("submissions/plots/frame_0931.jpg")
plot(i3)
gi3 <- image_ggplot(image_trim(i3))

i4 <- image_read("submissions/plots/frame_0958.jpg")
plot(i4)
gi4 <- image_ggplot(image_trim(i4))


Fig1A <- ggarrange(
  gi1,
  gi2,
  gi3,
  gi4,
  nrow = 1,
  align = "hv",
  labels = c("t = 0.00 s", "t = 0.04 s", "t = 0.32 s", "t = 1.08 s"),
  font.label = list(
    color = "white",
    face = "plain",
    size = 16
  ),
  label.y = 0.15,
  label.x = 0.3
)

Figure1agg <- ggarrange(
  Fig1A,
  Figure1,
  nrow = 2,
  labels = c("A", "B"),
  heights = c(1:2),
  font.label = list(size = 18)
)

Figure1agg

```

Save the figure 1

```{r Save Figure 1}
ggsave(
  filename = "Figure1.jpeg",
  plot = Figure1agg,
  device = "jpeg",
  path = "/Users/nicholaihensley/Desktop/Burglar Alarm/submissions/plots/",
  width = 16,
  height = 10,
  units = "in",
  dpi = 300
)
```

As a potential suggestion by reviewers and editors, we can perform the same analysis above but removing the outlier @ x = 5.

```{r Removing the high value outlier}


tb_2.1 <- msm_data_filtered %>%
  filter(
    food_sample == "photeros",
    treatment == "live",
    time_point < 4,
    num_luminescence < 5 | is.na(num_luminescence)
  ) %>%
  count(luminescence, attacked) %>%
  pivot_wider(names_from = attacked,
              values_from = n,
              values_fill = 0) %>%
  dplyr::select(-luminescence)

fisher.test(tb_2.1) # p-value = 7.093e-12

rate_cor_df_2.1 <- msm_data_filtered %>%
  group_by(exp_id) %>%
  mutate("was_eaten" = case_when(any(eaten == "yes") ~ "Eaten", .default = "Rejected")) %>%
  ungroup() %>%
  filter(food_sample == "photeros",
         treatment == "live",
         time_point < 4,
         attacked == "yes") %>%
  mutate(num_luminescence = case_when(is.na(num_luminescence) ~ 0, .default = num_luminescence)) %>%
  filter(num_luminescence < 5, genus != "Stegastes")

rate_model_2.1 <- glmmTMB(
  num_luminescence ~ num_attacks * genus + (1 | ind_id),
  family = poisson(),
  REML = FALSE,
  data = rate_cor_df_2.1,
  na.action = na.fail,
  #control = glmmTMBControl(optimizer = nlminb,optArgs = list(method = "Nelder-Mead"))
)

summary(rate_model_2.1)
check_model(rate_model_2.1)
check_overdispersion(rate_model_2.1)
check_distribution(rate_model_2.1)
check_zeroinflation(rate_model_1)
plot(effects::allEffects(rate_model_2.1))
sjPlot::tab_model(
  rate_model_2.1,
  show.intercept = FALSE,
  show.est = TRUE,
  show.dev = TRUE,
  show.loglik = TRUE,
  show.reflvl = TRUE
)

report::report_text(rate_model_2.1)

```

Plotting the supplementary analysis for Figure 1

```{r Plotting the supplementary Fig 1.1}

pred_grid <- rate_cor_df_2.1 %>% filter(genus != "Stegastes") %>%
  group_by(genus) %>%
  summarise(
    min_attacks = min(num_attacks, na.rm = TRUE),
    max_attacks = max(num_attacks, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  rowwise() %>%
  mutate(num_attacks = list(seq(min_attacks, max_attacks, length.out = 200))) %>%
  unnest(num_attacks) %>%
  mutate(genus = factor(genus, levels = levels(rate_cor_df$genus)))

pred <- predict(
  rate_model_2.1,
  newdata = pred_grid,
  type = "link",
  se.fit = TRUE,
  re.form = NA
)

plot_df <- pred_grid %>%
  mutate(
    fit_link = pred$fit,
    se_link = pred$se.fit,
    fit = exp(fit_link),
    lower = exp(fit_link - 1.96 * se_link),
    upper = exp(fit_link + 1.96 * se_link)
  )

FigureS2.1 <- rate_cor_df %>% filter(#was_eaten == "Rejected",
                                     num_luminescence < 5) %>%
  ggplot(aes(x = num_attacks, y = num_luminescence)) +
  geom_jitter(
    aes(
      shape = was_eaten,
      color = was_eaten,
      fill = was_eaten
    ),
    size = 5,
    stroke = 1.15,
    width = 0.05,
    height = 0.25
  ) +
  geom_jitter(
    data = rate_cor_df %>% filter(was_eaten == "Eaten", num_luminescence < 5),
    aes(
      shape = was_eaten,
      color = was_eaten,
      fill = was_eaten
    ),
    size = 5,
    stroke = 1.25,
    width = 0.05,
    height = 0.25
  ) +
  geom_line(
    data = plot_df %>% rename(num_luminescence = fit),
    size = 1,
    color = "#376795",
    aes(x = num_attacks, y = num_luminescence)
  ) +
  geom_ribbon(
    data =  plot_df %>% rename(num_luminescence = fit),
    aes(ymin = lower, ymax = upper),
    alpha = 0.25,
    fill = "#376795"
  ) +
  theme_bw(base_size = 20) +
  scale_shape_manual(name = "Final result", values = c("Eaten" = 3, "Rejected" = 21)) +
  scale_color_manual(name = "Final result",
                     values = c("Eaten" = "black", "Rejected" = "grey")) +
  scale_fill_manual(name = "Final result",
                    values = c("Eaten" = "black", "Rejected" = "white")) +
  xlab("Attack rate\n (fish bites per 20 s)") +
  ylab("Biuminescence rate\n (responses per 20 s)") + facet_grid(. ~ genus) +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "italic"))

FigureS2.1

```

Save the plot

```{r}

ggsave(
  filename = "FigureSr1.jpeg",
  plot = FigureS2.1,
  device = "jpeg",
  path = "/Users/nicholaihensley/Desktop/Burglar Alarm/submissions/plots/",
  width = 12,
  height = 8,
  units = "in",
  dpi = 300
)

```

Looking at the correlation between experience and time / trial

```{r Plotting the correlation between experience and time}

msm_data_filtered %>% #filter(genus != "Stegastes") %>%
  ggplot(aes(x = time_point, y = ind_id)) +
  geom_point(aes(size = luminescence_experience_c)) +
  facet_nested(relative_trial_c ~ treatment + type) +
  theme_minimal() + facet_grid(relative_trial ~ genus)

```

The genus Stegastes accumulates very little experience, so I will exclude it from the hazard and attack models downstream

## Looking at the Eaten Model

```{r Eaten Model}

hazard_model_1 <- glmmTMB(
  eaten ~ type + treatment_2 + relative_trial_c + luminescence_experience_c + order + time_point +
    genus +
    order:relative_trial_c +
    type:time_point +
    treatment_2:time_point +
    type:luminescence_experience_c +
    treatment_2:relative_trial_c +
    type:relative_trial_c +
    treatment_2:luminescence_experience_c +
    (1 | ind_id / trial / sample_id),
  family = binomial("cloglog"),
  REML = FALSE,
  data = msm_data_filtered %>% filter(genus != "Stegastes") ,
  #using the data set where at least one attack occurs in a trial
  na.action = na.fail,
  control = glmmTMBControl(optArgs = list(method = "L-BFGS-B"))
)

drop1(hazard_model_1, test = "Chisq")

```

Drop genus, order, order:trial, type:trial, and treatment:experience are not important without Stegastes in it, so will remove to be consistent with the attack rate model as well, as they don't experience much luminous responses

```{r Eaten Model Round 2}

hazard_model_2 <- glmmTMB(
  eaten ~ type + treatment_2 + relative_trial_c + luminescence_experience_c + time_point + order +
    treatment:time_point + # VIF too high to include
    type:luminescence_experience_c +
    treatment_2:relative_trial_c +
    (1 | ind_id / trial / sample_id),
  family = binomial("cloglog"),
  REML = FALSE,
  data = msm_data_filtered %>% filter(genus != "Stegastes"),
  #using the data set where at least one attack occurs in a trial
  na.action = na.fail,
  control = glmmTMBControl(
    optArgs = list(method = "L-BFGS-B"),
    optCtrl = list(iter.max = 500)
  )
)

drop1(hazard_model_2, test = "Chisq")

```


```{r Eaten Model Round 2}

hazard_model_2.25 <- glmmTMB(
  eaten ~ type + treatment_2 + relative_trial_c + luminescence_experience_c + time_point +
    #treatment:time_point + # VIF too high to include
    type:luminescence_experience_c +
    treatment_2:relative_trial_c +
    (1 | ind_id / trial / sample_id),
  family = binomial("cloglog"),
  REML = FALSE,
  data = msm_data_filtered %>% filter(genus != "Stegastes"),
  #using the data set where at least one attack occurs in a trial
  na.action = na.fail,
  control = glmmTMBControl(
    optArgs = list(method = "L-BFGS-B"),
    optCtrl = list(iter.max = 500)
  )
)

check_model(hazard_model_2.25)
check_collinearity(hazard_model_2.25)

```

Model wouldn't fit without removing the less important random effects. Time point interactions are too collinear to use, so dropping those. Now we can remove order.

```{r Eaten Model Comparison}

compare_performance(hazard_model_1, hazard_model_2.25, rank = TRUE)
plot(compare_performance(hazard_model_1, hazard_model_2.25, rank = TRUE))
test_performance(hazard_model_1, hazard_model_2.25)

```

Final model #2 is best for fixed effects. Now testing the random effects best.

```{r Randoem effects comparisons}

hazard_model_2.5 <- glmmTMB(
  eaten ~ type + treatment_2 + relative_trial_c + luminescence_experience_c + time_point +
    type:luminescence_experience_c +
    treatment_2:relative_trial_c +
    (1 | ind_id / trial / sample_id),
  family = binomial("cloglog"),
  REML = FALSE,
  data = msm_data_filtered %>% filter(genus != "Stegastes"),
  control = glmmTMBControl(optArgs = list(method = "L-BFGS-B"))
)

hazard_model_1.3 <- glmmTMB(
  eaten ~ type + treatment_2 + relative_trial_c + luminescence_experience_c + time_point + #genus +
    type:luminescence_experience_c +
    treatment_2:relative_trial_c +
    (1 | ind_id),
  family = binomial("cloglog"),
  REML = FALSE,
  data = msm_data_filtered %>% filter(genus != "Stegastes"),
  control = glmmTMBControl(optArgs = list(method = "L-BFGS-B"))
)

compare_performance(hazard_model_2.5, hazard_model_1.3)
plot(compare_performance(hazard_model_2.5, hazard_model_1.3, rank = TRUE))
test_performance(hazard_model_2.5, hazard_model_1.3)

```

Hazard model #3 has the best fitting random effects structure

```{r Eaten Model Results}


summary(hazard_model_1.3)
plot(effects::allEffects(hazard_model_1.3))
plot(effects::Effect(focal.predictors = "type", mod = hazard_model_1.3))
plot(effects::Effect(focal.predictors = "treatment_2", mod = hazard_model_1.3))
plot(effects::Effect(focal.predictors = "time_point", mod = hazard_model_1.3))
plot(
  effects::Effect(focal.predictors = "luminescence_experience_c", mod = hazard_model_1.3)
)
plot(effects::Effect(
  focal.predictors = c("relative_trial_c"),
  mod = hazard_model_1.3
))

car::Anova(hazard_model_1.3, "II")
check_model(hazard_model_1.3)
check_collinearity(hazard_model_1.3)
check_overdispersion(hazard_model_1.3)

sjPlot::plot_model(
  hazard_model_1.3,
  sort.est = "sort.all",
  pred.type = "fe",
  show.values = TRUE,
  value.offset = -0.35
)


sjPlot::tab_model(
  hazard_model_1.3,
  show.intercept = FALSE,
  show.est = TRUE,
  show.dev = TRUE,
  show.loglik = TRUE,
  show.reflvl = TRUE
)

plot(
  ggpredict(
    hazard_model_1.3,
    terms = ~ luminescence_experience_c + relative_trial_c + treatment_2 + type + time_point,
    typical = "marginalmeans",
    bias_correction = TRUE
  )
)

report::report_text(hazard_model_1.3)

```

Quickly plot the survival function across the time_points for simple visualization

```{r}


pred_grid <- expand.grid(
  type = levels(msm_data_filtered$type),
  treatment_2 = levels(msm_data_filtered$treatment_2),
  relative_trial_c = seq(1:4),
  luminescence_experience_c = 0,
  #seq(0,max(msm_data_filtered$luminescence_experience)),
  time_point = levels(msm_data_filtered$time_point)
)

pred_grid$time_point <- factor(
  pred_grid$time_point,
  levels = levels(msm_data_filtered$time_point),
  ordered = TRUE
)

X <- model.matrix(lme4::nobars(formula(hazard_model_1.3, component = "cond"))[-2], pred_grid)

beta_hat <- fixef(hazard_model_1.3)$cond
V_beta <- vcov(hazard_model_1.3)$cond

set.seed(123)

nsim <- 1000

beta_sim <- MASS::mvrnorm(nsim, mu = beta_hat, Sigma = V_beta)

eta_sim <- X %*% t(beta_sim)

haz_sim <- 1 - exp(-exp(eta_sim))

sim_df <- bind_cols(pred_grid, as.data.frame(haz_sim))

names(sim_df)[(ncol(pred_grid) + 1):ncol(sim_df)] <-
  paste0("sim_", seq_len(nsim))

plot_df <- sim_df %>%
  arrange(type, treatment_2, relative_trial_c, time_point) %>%
  group_by(type, treatment_2, relative_trial_c) %>%
  mutate(across(starts_with("sim_"), ~ 1 - cumprod(1 - .x))) %>%
  ungroup()

plot_df <- plot_df %>%
  rowwise() %>%
  mutate(
    prob = mean(c_across(starts_with("sim_"))),
    lower = quantile(c_across(starts_with("sim_")), 0.025),
    upper = quantile(c_across(starts_with("sim_")), 0.975)
  ) %>%
  ungroup() %>% mutate(type = case_when(type == "luminescent" ~ "bioluminescent",.default = type))

plot_df <- plot_df %>%
  mutate(condition = interaction(str_to_title(type), str_to_title(treatment_2)))

clean_figure <- ggplot(
  plot_df,
  aes(
    x = time_point,
    y = prob,
    colour = condition,
    fill = condition,
    group = condition
  )
) +
  geom_ribbon(aes(ymin = lower, ymax = upper),
              alpha = 0.15,
              colour = NA) +
  scale_color_manual(
    name = "Prey Type",
    breaks = c(
      'Bioluminescent.Live',
      'Nonluminescent.Live',
      'Bioluminescent.Anaesthetized',
      'Nonluminescent.Anaesthetized',
      'Bioluminescent.Dead',
      'Nonluminescent.Dead'
    ),
    values = c(
      'Bioluminescent.Live' = "#376795",
      "Nonluminescent.Live" = "#e76254",
      'Bioluminescent.Anaesthetized' = "#528fad",
      'Nonluminescent.Anaesthetized' = "#ef7a57",
      'Bioluminescent.Dead' = "#72bcd5",
      'Nonluminescent.Dead' = "#f7aa58"
    )
  ) +
  
  scale_fill_manual(
    name = "Prey Type",
    breaks = c(
      'Bioluminescent.Live',
      'Nonluminescent.Live',
      'Bioluminescent.Anaesthetized',
      'Nonluminescent.Anaesthetized',
      'Bioluminescent.Dead',
      'Nonluminescent.Dead'
    ),
    values = c(
      'Bioluminescent.Live' = "#376795",
      "Nonluminescent.Live" = "#e76254",
      'Bioluminescent.Anaesthetized' = "#528fad",
      'Nonluminescent.Anaesthetized' = "#ef7a57",
      'Bioluminescent.Dead' = "#72bcd5",
      'Nonluminescent.Dead' = "#f7aa58"
    )
  ) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  coord_cartesian(ylim = c(0, 1)) +
  labs(
    x = "Time point",
    y = "Cumulative probability of being eaten",
    colour = "Condition",
    fill = "Condition"
  ) +
  theme_bw(base_size = 20) +
  facet_nested(treatment_2 + type ~ relative_trial_c, nest_line = element_line()) +
  theme(strip.background = element_blank()) +
  labs(title = "Cumulative probability of being eaten over time and across trials", subtitle = "Separated by prey type and treatment")

ggsave(
  plot = clean_figure,
  filename = "./submissions/plots/cumulative_probability_eaten.jpeg",
  device = "jpeg",
  width = 15,
  height = 15,
  units = "in",
  dpi = 300,
  limitsize = TRUE
)

```

Will now generate the pairwise comparison table for the supplement, but comparing the relative risks for the different fixed effects

```{r}


## Pairwise comparisons for type and treatments shows in figure 2
emm_type_exp <- emmeans(
  hazard_model_1.3,
  ~ type + treatment_2 | luminescence_experience_c,
  at = list(
    luminescence_experience_c = 0,
    relative_trial_c = 0
  ),
  type = "response"
)

pairs(emm_type_exp)

## Pairwise slope comparisons for type shown in figure 3

exp_slopes <- emtrends(hazard_model_1.3, ~ type, var = "luminescence_experience_c")

summary(exp_slopes)
pairs(exp_slopes)

## Pairwise slope comparisons for treatment shown in figure 3

trial_slopes <- emtrends(hazard_model_1.3, ~ treatment_2, var = "relative_trial_c")

summary(trial_slopes)
pairs(trial_slopes)

```

Generate the prediction grid for hazard model and the survival function

```{r}
## ----------------------------
## 1. model-based survival predictions
## ----------------------------

pred_grid <- expand.grid(
  type = levels(msm_data_filtered$type),
  treatment_2 = levels(msm_data_filtered$treatment_2),
  relative_trial_c = 0,
  luminescence_experience_c = 0,
  time_point = levels(msm_data_filtered$time_point)
)

## model matrix for fixed effects
X <- model.matrix(delete.response(terms(hazard_model_1.3)), pred_grid)

## fixed-effect estimates and covariance matrix
beta_hat <- fixef(hazard_model_1.3)$cond
V_beta <- vcov(hazard_model_1.3)$cond

## simulate from fixed-effect uncertainty
set.seed(123)

nsim <- 1000

beta_sim <- MASS::mvrnorm(n = nsim, mu = beta_hat, Sigma = V_beta)

## convert linear predictor -> cloglog hazard
eta_sim <- X %*% t(beta_sim)

haz_sim <- 1 - exp(-exp(eta_sim))

## bind simulations to pred_grid
sim_df <- bind_cols(pred_grid, as.data.frame(haz_sim))

names(sim_df)[(ncol(pred_grid) + 1):ncol(sim_df)] <- paste0("sim_", seq_len(nsim))

## ----------------------------
## 2. hazards -> cumulative probability eaten within trial
## ----------------------------

plot_df <- sim_df %>%
  arrange(type, treatment_2, relative_trial_c, time_point) %>%
  group_by(type, treatment_2, relative_trial_c) %>%
  mutate(across(starts_with("sim_"), ~ 1 - cumprod(1 - .x))) %>%
  filter(time_point == levels(time_point)[nlevels(time_point)]) %>%
  ungroup() %>%
  group_by(type, treatment_2) %>%
  summarise(across(starts_with("sim_"), mean), .groups = "drop") %>%
  rowwise() %>%
  mutate(
    pred  = mean(c_across(starts_with("sim_"))),
    lower = quantile(c_across(starts_with("sim_")), 0.025),
    upper = quantile(c_across(starts_with("sim_")), 0.975)
  ) %>%
  ungroup() %>%
  mutate(
    type = str_to_title(type),
    treatment_2 = str_to_title(treatment_2),
    type = fct_relevel(type, c("Luminescent", "Nonluminescent")),
    type = case_when(type == "Luminescent" ~ "Bioluminescent",.default = type),
    type = fct_relevel(type, c("Bioluminescent", "Nonluminescent")),
    treatment_2 = fct_relevel(treatment_2, c("Live", "Anaesthetized", "Dead")),
    condition = interaction(type, treatment_2)
  )

## ----------------------------
## 3. pairwise contrasts on final cumulative probability eaten
## ----------------------------

sim_long <- plot_df %>%
  dplyr::select(type, treatment_2, condition, starts_with("sim_")) %>%
  tidyr::pivot_longer(cols = starts_with("sim_"),
                      names_to = "sim",
                      values_to = "pred_eaten")

contrast_df <- sim_long %>%
  dplyr::select(condition, sim, pred_eaten) %>%
  tidyr::pivot_wider(names_from = condition, values_from = pred_eaten)

pairwise_contrasts <- combn(levels(plot_df$condition), 2, simplify = FALSE) %>%
  purrr::map_dfr(function(x) {
    diff <- contrast_df[[x[1]]] - contrast_df[[x[2]]]
    
    tibble::tibble(
      group1 = x[1],
      group2 = x[2],
      estimate = mean(diff),
      lower = quantile(diff, 0.025),
      upper = quantile(diff, 0.975),
      significant = lower > 0 | upper < 0
    )
  })

library(multcompView)

p_mat <- matrix(
  1,
  nrow = nlevels(plot_df$condition),
  ncol = nlevels(plot_df$condition),
  dimnames = list(levels(plot_df$condition), levels(plot_df$condition))
)

for (i in seq_len(nrow(pairwise_contrasts))) {
  p_mat[pairwise_contrasts$group1[i], pairwise_contrasts$group2[i]] <-
    ifelse(pairwise_contrasts$significant[i], 0.01, 1)
  
  p_mat[pairwise_contrasts$group2[i], pairwise_contrasts$group1[i]] <-
    ifelse(pairwise_contrasts$significant[i], 0.01, 1)
}

letters_df <- multcompView::multcompLetters(p_mat, threshold = 0.05)$Letters %>%
  tibble::enframe(name = "condition", value = ".group")

plot_df <- plot_df %>%
  left_join(letters_df, by = "condition")

## visualizing the distributions of the simulations for each group
sim_long %>% ggplot(aes(x = pred_eaten)) +
  geom_histogram(aes(color = condition, fill = condition),
                 alpha = 0.15,
                 position = "identity") +
  theme_bw() + xlab("Cumulative Probability of Being Eaten (n = 1000)") + ylab("Count")

```

Now plotting the bar chart for figure 2

```{r Plotting the bar chat for Figure 2}


## ----------------------------
## 2. raw observed proportions for bars
## ----------------------------

bars_df <- msm_data_filtered %>% filter(genus != "Stegastes") %>%
  group_by(exp_id) %>%
  arrange(ind_id, trial, order, time_point, .by_group = TRUE) %>%
  filter(time_point == last(time_point)) %>%
  ungroup() %>%
  mutate(
    sample_id = case_when(
      str_detect(sample_id, "photeros.boiled") ~ "photeros.dead",
      str_detect(sample_id, "photeros.frozen") ~ "photeros.dead",
      str_detect(sample_id, "skogsbergia.boiled") ~ "skogsbergia.dead",
      str_detect(sample_id, "skogsbergia.frozen") ~ "skogsbergia.dead",
      TRUE ~ sample_id
    )
  ) %>%
  group_by(sample_id) %>%
  summarise(
    prop_eaten = mean(eaten == "yes"),
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(
    type = case_when(
      str_detect(sample_id, "photeros") ~ "Bioluminescent",
      str_detect(sample_id, "skogsbergia") ~ "Nonluminescent",
      TRUE ~ NA_character_
    ),
    treatment_2 = case_when(
      str_detect(sample_id, "live") ~ "Live",
      str_detect(sample_id, "anaesthetized") ~ "Anaesthetized",
      str_detect(sample_id, "dead") ~ "Dead",
      TRUE ~ NA_character_
    ),
    type = fct_relevel(type, c("Bioluminescent", "Nonluminescent")),
    treatment_2 = fct_relevel(treatment_2, c("Live", "Anaesthetized", "Dead")),
    condition = interaction(type, treatment_2)
  ) %>%
  filter(!is.na(type), !is.na(treatment_2))

## ----------------------------
## 3. plot
## ----------------------------

pd <- position_dodge2(width = 0.75)

Figure2 <- ggplot() +
  geom_col(
    data = bars_df,
    aes(
      x = treatment_2,
      y = prop_eaten,
      fill = condition,
      colour = condition
    ),
    position = pd,
    alpha = 1,
    width = 0.75,
    linewidth = 0.90
  ) +
  geom_text(
    data = bars_df,
    aes(
      x = treatment_2,
      y = -0.05,
      label = paste0("n = ", n),
      colour = condition
    ),
    position = pd,
    size = 5
  ) +
  geom_linerange(
    data = plot_df,
    aes(x = treatment_2, ymin = lower, ymax = upper),
    color = "black",
    position = pd,
    linewidth = 1.2
  ) +
  geom_point(
    data = plot_df,
    aes(
      x = treatment_2,
      y = pred,
      fill = condition,
      shape = condition
    ),
    color = "black",
    position = pd,
    size = 6.5,
    stroke = 1.2
  ) +
  geom_text(
    data = plot_df,
    aes(x = treatment_2, y = 1.05, label = .group),
    position = pd,
    size = 10,
    fontface = "bold"
  ) +
  scale_shape_manual(
    name = "Prey Type & Treatment",
    breaks = c(
      'Bioluminescent.Live',
      'Nonluminescent.Live',
      'Bioluminescent.Anaesthetized',
      'Nonluminescent.Anaesthetized',
      'Bioluminescent.Dead',
      'Nonluminescent.Dead'
    ),
    values = c(
      'Bioluminescent.Live' = 25,
      "Nonluminescent.Live" = 24,
      'Bioluminescent.Anaesthetized' = 25,
      'Nonluminescent.Anaesthetized' = 24,
      'Bioluminescent.Dead' = 25,
      'Nonluminescent.Dead' = 24
    )
  ) +
  scale_color_manual(
    name = "Prey Type & Treatment",
    breaks = c(
      'Bioluminescent.Live',
      'Nonluminescent.Live',
      'Bioluminescent.Anaesthetized',
      'Nonluminescent.Anaesthetized',
      'Bioluminescent.Dead',
      'Nonluminescent.Dead'
    ),
    values = c(
      'Bioluminescent.Live' = "#376795",
      "Nonluminescent.Live" = "#e76254",
      'Bioluminescent.Anaesthetized' = "#528fad",
      'Nonluminescent.Anaesthetized' = "#ef7a57",
      'Bioluminescent.Dead' = "#72bcd5",
      'Nonluminescent.Dead' = "#f7aa58"
    )
  ) +
  
  scale_fill_manual(
    name = "Prey Type & Treatment",
    breaks = c(
      'Bioluminescent.Live',
      'Nonluminescent.Live',
      'Bioluminescent.Anaesthetized',
      'Nonluminescent.Anaesthetized',
      'Bioluminescent.Dead',
      'Nonluminescent.Dead'
    ),
    values = c(
      'Bioluminescent.Live' = "#376795",
      "Nonluminescent.Live" = "#e76254",
      'Bioluminescent.Anaesthetized' = "#528fad",
      'Nonluminescent.Anaesthetized' = "#ef7a57",
      'Bioluminescent.Dead' = "#72bcd5",
      'Nonluminescent.Dead' = "#f7aa58"
    )
  ) +
  coord_cartesian(ylim = c(-0.07, 1.1)) +
  scale_y_continuous(breaks = c(0.25, 0.50, 0.75, 1)) +
  labs(x = "Treatment", y = "Proportion eaten (bars)\nCumulative predicted probability of being eaten + 95% CI (triangles + whiskers)") +
  theme_bw(base_size = 20) + theme(legend.position = c(0.2,0.7))

Figure2

```


```{r Saving figure2}

ggsave(
  plot = Figure2,
  filename = "Figure2.jpeg",
  device = "jpeg",
  width = 14,
  height = 11,
  units = "in",
  dpi = 300,
  path = "/Users/nicholaihensley/Desktop/Burglar Alarm/submissions/plots/"
)

```

Plotting the hazard model other fixed effects

```{r Plotting the Eaten hazard model}


library(MASS)
library(dplyr)
library(tidyr)
library(stringr)
library(forcats)
library(ggplot2)

inv_cloglog <- function(eta) {
  1 - exp(-exp(eta))
}

lum_grid <- seq(
  min(msm_data_filtered$luminescence_experience, na.rm = TRUE),
  max(msm_data_filtered$luminescence_experience, na.rm = TRUE),
  length.out = 200
)

trial_seq <- sort(unique(msm_data_filtered$relative_trial_c))

lum_pred_grid <- expand.grid(
  type = levels(msm_data_filtered$type),
  treatment_2 = levels(msm_data_filtered$treatment_2),
  luminescence_experience_c = lum_grid,
  relative_trial_c = 0,
  time_point = levels(msm_data_filtered$time_point)
)

X <- model.matrix(lme4::nobars(formula(hazard_model_1.3, component = "cond"))[-2], lum_pred_grid)

beta_hat <- fixef(hazard_model_1.3)$cond
V_beta <- vcov(hazard_model_1.3)$cond

set.seed(123)
nsim <- 1000

beta_sim <- MASS::mvrnorm(nsim, beta_hat, V_beta)

eta_sim <- X %*% t(beta_sim)
haz_sim <- inv_cloglog(eta_sim)

sim_df <- bind_cols(lum_pred_grid, as.data.frame(haz_sim))

names(sim_df)[(ncol(lum_pred_grid) + 1):ncol(sim_df)] <- paste0("sim_", seq_len(nsim))

lum_df <- sim_df %>%
  arrange(type,
          treatment_2,
          luminescence_experience_c,
          relative_trial_c,
          time_point) %>%
  group_by(type,
           treatment_2,
           luminescence_experience_c,
           relative_trial_c) %>%
  mutate(across(starts_with("sim_"), ~ 1 - cumprod(1 - .x))) %>%
  filter(time_point == levels(time_point)[nlevels(time_point)]) %>%
  ungroup() %>%
  group_by(type, luminescence_experience_c) %>%
  summarise(across(starts_with("sim_"), mean), .groups = "drop") %>%
  rowwise() %>%
  mutate(
    prob = mean(c_across(starts_with("sim_"))),
    asymp.LCL = quantile(c_across(starts_with("sim_")), 0.025),
    asymp.UCL = quantile(c_across(starts_with("sim_")), 0.975)
  ) %>%
  ungroup() %>%
  mutate(x = luminescence_experience_c, type = str_to_title(type), 
         type = case_when(type == "Luminescent" ~ "Bioluminescent",.default = type))

Fig3.p1 <- ggplot(lum_df, aes(
  x = x,
  y = prob,
  colour = type,
  fill = type
)) +
  geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL),
              alpha = 0.15,
              colour = NA) +
  geom_line(linewidth = 1) +
  geom_rug(
    data = subset(msm_data_filtered %>% mutate(type = str_to_title(type)), eaten == "no"),
    aes(x = luminescence_experience, y = 0, color = type),
    position = position_jitter(width = 0.05, height = 0),
    inherit.aes = FALSE,
    sides = "b",
    alpha = 0.25
  ) +
  geom_rug(
    data = subset(
      msm_data_filtered %>% mutate(type = str_to_title(type)),
      eaten == "yes"
    ),
    aes(x = luminescence_experience, y = 1, color = type),
    position = position_jitter(width = 0.05, height = 0),
    inherit.aes = FALSE,
    sides = "t",
    alpha = 0.25,
  ) +
  scale_color_manual(
    name = "Prey Type",
    values = c(
      "Bioluminescent" = "#376795",
      "Nonluminescent" = "#e76254"
    )
  ) +
  scale_fill_manual(
    name = "Prey Type",
    values = c(
      "Bioluminescent" = "#376795",
      "Nonluminescent" = "#e76254"
    )
  ) +
  labs(
    x = "Bioluminescence experience",
    y = "Cumulative predicted probability of being eaten",
    colour = "Prey Type",
    fill = "Prey Type"
  ) +
  coord_cartesian(ylim = c(0, 1)) +
  theme_bw(base_size = 20) +
  theme(legend.position = "top")

Fig3.p1

ggsave(
  plot = Fig3.p1,
  filename = "Fig3.p1.jpeg",
  device = "jpeg",
  width = 10,
  height = 8,
  units = "in",
  dpi = 300,
  path = "/Users/nicholaihensley/Desktop/Burglar Alarm/submissions/plots/"
)

```


```{r}

trial_grid <- seq(
  min(msm_data_filtered$relative_trial, na.rm = TRUE),
  max(msm_data_filtered$relative_trial, na.rm = TRUE),
  length.out = 200
)

trial_pred_grid <- expand.grid(
  type = levels(msm_data_filtered$type),
  treatment_2 = levels(msm_data_filtered$treatment_2),
  relative_trial_c = trial_grid,
  luminescence_experience_c = 0,
  time_point = levels(msm_data_filtered$time_point)
)

X <- model.matrix(lme4::nobars(formula(hazard_model_1.3, component = "cond"))[-2], trial_pred_grid)

eta_sim <- X %*% t(beta_sim)
haz_sim <- inv_cloglog(eta_sim)

sim_df <- bind_cols(trial_pred_grid, as.data.frame(haz_sim))

names(sim_df)[(ncol(trial_pred_grid) + 1):ncol(sim_df)] <- paste0("sim_", seq_len(nsim))

trial_df <- sim_df %>%
  arrange(treatment_2,
          type,
          luminescence_experience_c,
          relative_trial_c,
          time_point) %>%
  group_by(treatment_2,
           type,
           luminescence_experience_c,
           relative_trial_c) %>%
  mutate(across(starts_with("sim_"), ~ 1 - cumprod(1 - .x))) %>%
  filter(time_point == levels(time_point)[nlevels(time_point)]) %>%
  ungroup() %>%
  group_by(treatment_2, relative_trial_c) %>%
  summarise(across(starts_with("sim_"), mean), .groups = "drop") %>%
  rowwise() %>%
  mutate(
    prob = mean(c_across(starts_with("sim_"))),
    asymp.LCL = quantile(c_across(starts_with("sim_")), 0.025),
    asymp.UCL = quantile(c_across(starts_with("sim_")), 0.975)
  ) %>%
  ungroup() %>%
  mutate(
    x = relative_trial_c,
    treatment_2 = str_to_title(treatment_2),
    treatment_2 = fct_relevel(treatment_2, c("Live", "Anaesthetized", "Dead"))
  )

Fig3.p2 <- ggplot(trial_df,
                  aes(
                    x = x,
                    y = prob,
                    colour = treatment_2,
                    fill = treatment_2
                  )) +
  geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL),
              alpha = 0.15,
              colour = NA) +
  geom_line(linewidth = 1) +
  geom_rug(
    data = subset(
      msm_data_filtered %>% mutate(treatment_2 = str_to_title(treatment_2)),
      eaten == "no"
    ),
    aes(x = relative_trial, y = 0, color = treatment_2),
    position = position_jitter(width = 0.15, height = 0),
    inherit.aes = FALSE,
    sides = "b",
    alpha = 0.25
  ) +
  geom_rug(
    data = subset(
      msm_data_filtered %>% mutate(treatment_2 = str_to_title(treatment_2)),
      eaten == "yes"
    ),
    aes(x = relative_trial, y = 1, color = treatment_2),
    position = position_jitter(width = 0.15, height = 0),
    inherit.aes = FALSE,
    sides = "t",
    alpha = 0.25,
  ) +
  scale_color_manual(
    name = "Prey Treatment",
    values = c(
      "Live" = "#302F8AFF",
      "Anaesthetized" = "#602F8AFF",
      "Dead" = "#9B4F82FF"
    )
    )+
  scale_fill_manual(
    name = "Prey Treatment",
    values = c(
      "Live" = "#302F8AFF",
      "Anaesthetized" = "#602F8AFF",
      "Dead" = "#9B4F82FF"
    )
  ) +
  labs(
    x = "Trial number",
    y = "Cumulative predicted probability of being eaten",
    colour = "Prey Treatment",
    fill = "Prey Treatment"
  ) +
  coord_cartesian(ylim = c(0, 1)) +
  theme_bw(base_size = 20) +
  theme(legend.position = "top")

Fig3.p2

ggsave(
  plot = Fig3.p2,
  filename = "Fig3.p2.jpeg",
  device = "jpeg",
  width = 10,
  height = 8,
  units = "in",
  dpi = 300,
  path = "/Users/nicholaihensley/Desktop/Burglar Alarm/submissions/plots/"
)

```

Now plotting Figure 3

```{r}

# Combinding the fig3 plots
Fig3 <- ggarrange(
  plotlist = c(Fig3.p1, Fig3.p2),
  ncol = 2,
  labels = "AUTO",
  align = "hv",
  font.label = list(size = 20)
)

ggsave(
  plot = Fig3.p1,
  filename = "Figure3.jpeg",
  device = "jpeg",
  width = 9,
  height = 8,
  units = "in",
  dpi = 300,
  path = "/Users/nicholaihensley/Desktop/Burglar Alarm/submissions/plots/"
)

ggsave(
  plot = Fig3.p2,
  filename = "Figure3p2_supp.jpeg",
  device = "jpeg",
  width = 9,
  height = 8,
  units = "in",
  dpi = 300,
  path = "/Users/nicholaihensley/Desktop/Burglar Alarm/submissions/plots/"
)

```

## Supplementary Results

Plotting all the data even positive controls for the proportion eaten of each type and treatment

```{r Plotting all eaten data for the supplement}

bars_all_df <- data_filtered %>%
  filter(
    food_sample != "worm" |
      food_sample != "cumacean" |
      food_sample != "mysid",
    genus != "Stegastes"
  ) %>%
  group_by(exp_id) %>%
  arrange(ind_id, trial, order, time_point, .by_group = TRUE) %>%
  slice_max(time_point) %>%
  ungroup() %>%
  mutate(
    sample_id = case_when(
      str_detect(sample_id, "photeros.boiled") ~ "photeros.dead",
      str_detect(sample_id, "photeros.frozen") ~ "photeros.dead",
      str_detect(sample_id, "skogsbergia.boiled") ~ "skogsbergia.dead",
      str_detect(sample_id, "skogsbergia.frozen") ~ "skogsbergia.dead",
      str_detect(sample_id, "isopod.control") ~ "isopod.control",
      str_detect(sample_id, "fish.control") ~ "fish.control",
      TRUE ~ sample_id
    )
  ) %>%
  group_by(sample_id) %>%
  summarise(
    prop_eaten = mean(eaten == "yes"),
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(
    type = case_when(
      str_detect(sample_id, "photeros") ~ "Bioluminescent",
      str_detect(sample_id, "skogsbergia") ~ "Nonluminescent",
      str_detect(sample_id, "isopod") ~ "Isopod",
      str_detect(sample_id, "fish") ~ "Fish",
      TRUE ~ NA_character_
    ),
    treatment_2 = case_when(
      str_detect(sample_id, "live") ~ "Live",
      str_detect(sample_id, "anaesthetized") ~ "Anaesthetized",
      str_detect(sample_id, "dead") ~ "Dead",
      str_detect(sample_id, "control") ~ "Control",
      TRUE ~ NA_character_
    ),
    type = fct_relevel(type, c(
      "Fish", "Isopod", "Bioluminescent", "Nonluminescent"
    )),
    treatment_2 = fct_relevel(treatment_2, c(
      "Control", "Live", "Anaesthetized", "Dead"
    )),
    condition = interaction(type, treatment_2)
  ) %>%
  filter(!is.na(type), !is.na(treatment_2))

FigureS5 <- ggplot() +
  geom_col(
    data = bars_all_df,
    aes(
      x = treatment_2,
      y = prop_eaten,
      fill = condition,
      colour = condition
    ),
    position = pd,
    alpha = 1,
    width = 0.75,
    linewidth = 0.90
  ) +
  geom_text(
    data = bars_all_df,
    aes(
      x = treatment_2,
      y = -0.05,
      label = paste0("n = ", n),
      colour = condition
    ),
    position = pd,
    size = 5
  ) +
  scale_color_manual(
    name = "Prey Type",
    breaks = c(
      'Fish.Control',
      'Isopod.Control',
      'Bioluminescent.Live',
      'Nonluminescent.Live',
      'Bioluminescent.Anaesthetized',
      'Nonluminescent.Anaesthetized',
      'Bioluminescent.Dead',
      'Nonluminescent.Dead'
    ),
    values = c(
      "Fish.Control" = "grey55",
      "Isopod.Control" = "grey75",
      'Bioluminescent.Live' = "#376795",
      "Nonluminescent.Live" = "#e76254",
      'Bioluminescent.Anaesthetized' = "#528fad",
      'Nonluminescent.Anaesthetized' = "#ef7a57",
      'Bioluminescent.Dead' = "#72bcd5",
      'Nonluminescent.Dead' = "#f7aa58"
    )
  ) +
  
  scale_fill_manual(
    name = "Prey Type",
    breaks = c(
      'Fish.Control',
      'Isopod.Control',
      'Bioluminescent.Live',
      'Nonluminescent.Live',
      'Bioluminescent.Anaesthetized',
      'Nonluminescent.Anaesthetized',
      'Bioluminescent.Dead',
      'Nonluminescent.Dead'
    ),
    values = c(
      "Fish.Control" = "grey55",
      "Isopod.Control" = "grey75",
      'Bioluminescent.Live' = "#376795",
      "Nonluminescent.Live" = "#e76254",
      'Bioluminescent.Anaesthetized' = "#528fad",
      'Nonluminescent.Anaesthetized' = "#ef7a57",
      'Bioluminescent.Dead' = "#72bcd5",
      'Nonluminescent.Dead' = "#f7aa58"
    )
  ) +
  
  coord_cartesian(ylim = c(-0.07, 1.1)) +
  scale_y_continuous(breaks = c(0.25, 0.50, 0.75, 1)) +
  labs(
    title = "Proportion of prey eaten",
    subtitle = "Calculated as: n eaten / N presentations with attack \u2265 1",
    x = "Treatment",
    y = "Proportion eaten (bars)"
  ) +
  theme_bw(base_size = 20)

FigureS5

ggsave(
  plot = FigureS5,
  filename = "FigureS5.jpeg",
  device = "jpeg",
  width = 13,
  height = 8,
  units = "in",
  dpi = 300,
  path = "/Users/nicholaihensley/Desktop/Burglar Alarm/submissions/plots/"
)

```

### Running the hazard model with the dead treatments separated for the supplement

```{r Frozen and Boiled as separate effects}


hazard_model_s1.1 <- glmmTMB(
  eaten ~ type + treatment + relative_trial_c + luminescence_experience_c + 
    order + time_point + genus + order:relative_trial_c +
    type:time_point +
    treatment:time_point +
    type:luminescence_experience_c +
    treatment:relative_trial_c +
    type:relative_trial_c +
    treatment:luminescence_experience_c +
    (1 | ind_id / trial / sample_id),
  family = binomial("cloglog"),
  REML = FALSE,
  data = msm_data_filtered %>% filter(genus != "Stegastes"),
  #using the data set where at least one attack occurs in a trial
  na.action = na.fail,
  control = glmmTMBControl(optArgs = list(method = "L-BFGS-B"))
)

drop1(hazard_model_s1.1, test = "Chisq")

```

Same results as in the main body of the article

```{r Variable reduction}

hazard_model_s1.2 <- glmmTMB(
  eaten ~ type + treatment + relative_trial_c + luminescence_experience_c + time_point +
    # treatment:time_point + # VIF too high to include
    type:luminescence_experience_c +
    treatment:relative_trial_c +
    (1 | ind_id / trial / sample_id),
  family = binomial("cloglog"),
  REML = FALSE,
  data = msm_data_filtered %>% filter(genus != "Stegastes"),
  #using the data set where at least one attack occurs in a trial
  na.action = na.fail,
  control = glmmTMBControl(
    optArgs = list(method = "L-BFGS-B"),
    optCtrl = list(iter.max = 500)
  )
)

check_model(hazard_model_s1.2)
check_collinearity(hazard_model_s1.2)

compare_performance(hazard_model_s1.1, hazard_model_s1.2, rank = TRUE)
plot(compare_performance(hazard_model_s1.1, hazard_model_s1.2, rank = TRUE))
test_performance(hazard_model_s1.1, hazard_model_s1.2)

### model s1.2 is better, just as in the regular data
```

Same as the main model as well

```{r Looking at the random effects of the new hazard model}

hazard_model_s1.25 <- glmmTMB(
  eaten ~ type + treatment + relative_trial_c + luminescence_experience_c + time_point +
    type:luminescence_experience_c +
    treatment:relative_trial_c +
    (1 | ind_id / trial / sample_id),
  family = binomial("cloglog"),
  REML = FALSE,
  data = msm_data_filtered %>% filter(genus != "Stegastes"),
  control = glmmTMBControl(optArgs = list(method = "L-BFGS-B"))
)

hazard_model_s1.3 <- glmmTMB(
  eaten ~ type + treatment + relative_trial_c + luminescence_experience_c + time_point +
    type:luminescence_experience_c +
    treatment:relative_trial_c +
    (1 | ind_id),
  family = binomial("cloglog"),
  REML = FALSE,
  data = msm_data_filtered %>% filter(genus != "Stegastes"),
  control = glmmTMBControl(optArgs = list(method = "L-BFGS-B"))
)

compare_performance(hazard_model_s1.25, hazard_model_s1.3)
plot(compare_performance(hazard_model_s1.25, hazard_model_s1.3, rank = TRUE))
test_performance(hazard_model_s1.25, hazard_model_s1.3)

```

Hazard model s1.3 has the best fitting random effects structure

```{r New Eaten Model Results}

summary(hazard_model_s1.3)
plot(effects::allEffects(hazard_model_s1.3))
plot(effects::Effect(focal.predictors = "type", mod = hazard_model_s1.3))
plot(effects::Effect(focal.predictors = "treatment", mod = hazard_model_s1.3))
plot(effects::Effect(focal.predictors = "time_point", mod = hazard_model_s1.3))
plot(
  effects::Effect(focal.predictors = "luminescence_experience_c", mod = hazard_model_s1.3)
)
plot(effects::Effect(
  focal.predictors = c("relative_trial_c"),
  mod = hazard_model_s1.3
))

car::Anova(hazard_model_s1.3, "II")
check_model(hazard_model_s1.3)
check_collinearity(hazard_model_s1.3)
check_overdispersion(hazard_model_s1.3)

sjPlot::plot_model(
  hazard_model_s1.3,
  sort.est = "sort.all",
  pred.type = "fe",
  show.values = TRUE,
  value.offset = -0.35
)


sjPlot::tab_model(
  hazard_model_s1.3,
  show.intercept = FALSE,
  show.est = TRUE,
  show.dev = TRUE,
  show.loglik = TRUE,
  show.reflvl = TRUE
)

plot(
  ggpredict(
    hazard_model_s1.3,
    terms = ~ luminescence_experience_c + relative_trial_c + treatment + type + time_point,
    typical = "marginalmeans",
    bias_correction = TRUE
  )
)

```


```{r Pairwise comparisons}


## Pairwise comparisons for type and treatments shows in figure 2
emm_type_exp <- emmeans(
  hazard_model_s1.3,
  ~ type + treatment | luminescence_experience_c,
  at = list(
    luminescence_experience_c = 0,
    relative_trial_c = 0
  ),
  type = "response"
)

pairs(emm_type_exp)

## Pairwise slope comparisons for type shown in figure 3

exp_slopes <- emtrends(hazard_model_s1.3, ~ type, var = "luminescence_experience_c")

summary(exp_slopes)
pairs(exp_slopes)

## Pairwise slope comparisons for treatment shown in figure 3

trial_slopes <- emtrends(hazard_model_s1.3, ~ treatment, var = "relative_trial_c")

summary(trial_slopes)
pairs(trial_slopes)

```

Generate the prediction grid for hazard model and the survival function

```{r New eaten model expanded results}

## ----------------------------
## 1. model-based survival predictions
## ----------------------------

pred_grid <- expand.grid(
  type = levels(msm_data_filtered$type),
  treatment = levels(msm_data_filtered$treatment),
  relative_trial_c = 0,
  luminescence_experience_c = 0,
  time_point = levels(msm_data_filtered$time_point)
)

## model matrix for fixed effects
X <- model.matrix(delete.response(terms(hazard_model_s1.3)), pred_grid)

## fixed-effect estimates and covariance matrix
beta_hat <- fixef(hazard_model_s1.3)$cond
V_beta <- vcov(hazard_model_s1.3)$cond

## simulate from fixed-effect uncertainty
set.seed(123)

nsim <- 1000

beta_sim <- MASS::mvrnorm(n = nsim, mu = beta_hat, Sigma = V_beta)

## convert linear predictor -> cloglog hazard
eta_sim <- X %*% t(beta_sim)

haz_sim <- 1 - exp(-exp(eta_sim))

## bind simulations to pred_grid
sim_df <- bind_cols(pred_grid, as.data.frame(haz_sim))

names(sim_df)[(ncol(pred_grid) + 1):ncol(sim_df)] <- paste0("sim_", seq_len(nsim))

## ----------------------------
## 2. hazards -> cumulative probability eaten within trial
## ----------------------------

plot_df <- sim_df %>%
  arrange(type, treatment, relative_trial_c, time_point) %>%
  group_by(type, treatment, relative_trial_c) %>%
  mutate(across(starts_with("sim_"), ~ 1 - cumprod(1 - .x))) %>%
  filter(time_point == levels(time_point)[nlevels(time_point)]) %>%
  ungroup() %>%
  group_by(type, treatment) %>%
  summarise(across(starts_with("sim_"), mean), .groups = "drop") %>%
  rowwise() %>%
  mutate(
    pred  = mean(c_across(starts_with("sim_"))),
    lower = quantile(c_across(starts_with("sim_")), 0.025),
    upper = quantile(c_across(starts_with("sim_")), 0.975)
  ) %>%
  ungroup() %>%
  mutate(
    type = str_to_title(type),
    type = case_when(type == "Luminescent" ~ "Bioluminescent",.default = type),
    treatment = str_to_title(treatment),
    type = fct_relevel(type, c("Bioluminescent", "Nonluminescent")),
    treatment = fct_relevel(treatment, c(
      "Live", "Anaesthetized", "Frozen", "Boiled"
    )),
    condition = interaction(type, treatment)
  )

## ----------------------------
## 3. pairwise contrasts on final cumulative probability eaten
## ----------------------------

sim_long <- plot_df %>%
  dplyr::select(type, treatment, condition, starts_with("sim_")) %>%
  tidyr::pivot_longer(cols = starts_with("sim_"),
                      names_to = "sim",
                      values_to = "pred_eaten")

contrast_df <- sim_long %>%
  dplyr::select(condition, sim, pred_eaten) %>%
  tidyr::pivot_wider(names_from = condition, values_from = pred_eaten)

pairwise_contrasts <- combn(levels(plot_df$condition), 2, simplify = FALSE) %>%
  purrr::map_dfr(function(x) {
    diff <- contrast_df[[x[1]]] - contrast_df[[x[2]]]
    
    tibble::tibble(
      group1 = x[1],
      group2 = x[2],
      estimate = mean(diff),
      lower = quantile(diff, 0.025),
      upper = quantile(diff, 0.975),
      significant = lower > 0 | upper < 0
    )
  })

library(multcompView)

p_mat <- matrix(
  1,
  nrow = nlevels(plot_df$condition),
  ncol = nlevels(plot_df$condition),
  dimnames = list(levels(plot_df$condition), levels(plot_df$condition))
)

for (i in seq_len(nrow(pairwise_contrasts))) {
  p_mat[pairwise_contrasts$group1[i], pairwise_contrasts$group2[i]] <-
    ifelse(pairwise_contrasts$significant[i], 0.01, 1)
  
  p_mat[pairwise_contrasts$group2[i], pairwise_contrasts$group1[i]] <-
    ifelse(pairwise_contrasts$significant[i], 0.01, 1)
}

letters_df <- multcompView::multcompLetters(p_mat, threshold = 0.05)$Letters %>%
  tibble::enframe(name = "condition", value = ".group")

plot_df <- plot_df %>%
  left_join(letters_df, by = "condition")

## visualizing the distributions of the simulations for each group
sim_long %>% ggplot(aes(x = pred_eaten)) +
  geom_histogram(aes(color = condition, fill = condition),
                 alpha = 0.15,
                 position = "identity") +
  theme_bw() + xlab("Cumulative Probability of Being Eaten (n = 1000)") + ylab("Count")
```

Now plotting the bar chart for supplementary figure 2

```{r Plotting the bar chat for Figure 2}

## ----------------------------
## 2. raw observed proportions for bars
## ----------------------------

bars_df <- msm_data_filtered %>% filter(genus != "Stegastes") %>% 
  group_by(exp_id) %>%
  arrange(ind_id, trial, order, time_point, .by_group = TRUE) %>%
  filter(time_point == last(time_point)) %>%
  ungroup() %>% 
  group_by(sample_id) %>%
  summarise(
    prop_eaten = mean(eaten == "yes"),
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(
    type = case_when(
      str_detect(sample_id, "photeros") ~ "Bioluminescent",
      str_detect(sample_id, "skogsbergia") ~ "Nonluminescent",
      TRUE ~ NA_character_
    ),
    treatment = case_when(
      str_detect(sample_id, "live") ~ "Live",
      str_detect(sample_id, "anaesthetized") ~ "Anaesthetized",
      str_detect(sample_id, "frozen") ~ "Frozen",
      str_detect(sample_id, "boiled") ~ "Boiled",
      TRUE ~ NA_character_
    ),
    type = fct_relevel(type, c("Bioluminescent", "Nonluminescent")),
    treatment = fct_relevel(treatment, c("Live", "Anaesthetized", "Frozen", "Boiled")),
    condition = interaction(type,treatment)
  ) %>%
  filter(!is.na(type), !is.na(treatment))

## ----------------------------
## 3. plot
## ----------------------------

pd <- position_dodge2(width = 0.75)

Figure2.1 <- ggplot() +
  geom_col(
    data = bars_df,
    aes(x = treatment, y = prop_eaten, fill = condition, colour = condition),
    position = pd,
    alpha = 1,
    width = 0.75,
    linewidth = 0.90
  ) +
  geom_text(
    data = bars_df,
    aes(x = treatment, y = -0.05, label = paste0("n = ", n), colour = condition),
    position = pd,
    size = 5
  ) +
  geom_linerange(
    data = plot_df,
    aes(x = treatment, ymin = lower, ymax = upper),
    color = "black",
    position = pd,
    linewidth = 1.2
  ) +
  geom_point(
    data = plot_df,
    aes(x = treatment, y = pred, fill = condition, shape = condition),
    color = "black",
    position = pd,
    size = 6.5,
    stroke = 1.2
  ) +
  geom_text(
    data = plot_df,
    aes(x = treatment, y = 1.05, label = .group),
    position = pd,
    size = 10,
    fontface = "bold"
  ) +

  scale_shape_manual(name = "Prey Type",
                     breaks = c('Bioluminescent.Live',
                                'Nonluminescent.Live',
                                'Bioluminescent.Anaesthetized',
                                'Nonluminescent.Anaesthetized',
                                'Bioluminescent.Frozen',
                                'Nonluminescent.Frozen',
                                'Bioluminescent.Boiled',
                                'Nonluminescent.Boiled'
                                ),
                     values = c('Bioluminescent.Live' = 25,
                                "Nonluminescent.Live" = 24,
                                'Bioluminescent.Anaesthetized' = 25,
                                'Nonluminescent.Anaesthetized' = 24,
                                'Bioluminescent.Frozen' = 25,
                                'Nonluminescent.Frozen' = 24,
                                'Bioluminescent.Boiled' = 25,
                                'Nonluminescent.Boiled' = 24
                                )
  ) +
  
  scale_color_manual(name = "Prey Type",
                      breaks = c('Bioluminescent.Live',
                                'Nonluminescent.Live',
                                'Bioluminescent.Anaesthetized',
                                'Nonluminescent.Anaesthetized',
                                'Bioluminescent.Frozen',
                                'Nonluminescent.Frozen',
                                'Bioluminescent.Boiled',
                                'Nonluminescent.Boiled'
                                ),
                     values = c('Bioluminescent.Live' = "#376795",
                                "Nonluminescent.Live" = "#e76254",
                                'Bioluminescent.Anaesthetized' = "#528fad",
                                'Nonluminescent.Anaesthetized' = "#ef7a57",
                                'Bioluminescent.Frozen' = "#71bcd5",
                                'Bioluminescent.Boiled' = "#91bee9",
                                'Nonluminescent.Frozen' = "#f7aa58",
                                'Nonluminescent.Boiled' = "#f7cc58"
                                )
  ) +
  
  scale_fill_manual(name = "Prey Type",
                      breaks = c('Bioluminescent.Live',
                                'Nonluminescent.Live',
                                'Bioluminescent.Anaesthetized',
                                'Nonluminescent.Anaesthetized',
                                'Bioluminescent.Frozen',
                                'Nonluminescent.Frozen',
                                'Bioluminescent.Boiled',
                                'Nonluminescent.Boiled'
                                ),
                     values = c('Bioluminescent.Live' = "#376795",
                                "Nonluminescent.Live" = "#e76254",
                                'Bioluminescent.Anaesthetized' = "#528fad",
                                'Nonluminescent.Anaesthetized' = "#ef7a57",
                                'Bioluminescent.Frozen' = "#71bcd5",
                                'Bioluminescent.Boiled' = "#91bee9",
                                'Nonluminescent.Frozen' = "#f7aa58",
                                'Nonluminescent.Boiled' = "#f7cc58"
                                )
  ) +
  
  coord_cartesian(ylim = c(-0.07, 1.1)) +
  scale_y_continuous(breaks = c(0.25,0.50,0.75,1)) +
  labs(
    x = "Treatment",
    y = "Proportion eaten (bars)\nCumulative predicted probability of being eaten + 95% CI (triangles + whiskers)"
  ) +
  theme_bw(base_size = 20)

Figure2.1

ggsave(filename = "FigureS2.1.jpeg",plot = Figure2.1,device = "jpeg",
       path = "./submissions/plots/",width = 18,height = 12,units = "in",dpi = 300)

```


### A separate analysis looking at experience with Photeros during attacks versus luminous signal exposure

```{r Photeros experience model}


msm_photeros <- msm_data_filtered %>%
  filter(genus != "Stegastes") %>%
  tidyr::replace_na(list(photeros_attack_experience = 0))

## same as the hazard_model_1.3 but using the redefined dataset to use the photeros experience without NAs
hazard_model_4 <- glmmTMB(
  eaten ~ type + treatment_2 + relative_trial_c +
    luminescence_experience_c +
    # photeros_attack_experience +
    time_point +
    type:luminescence_experience_c +
    treatment_2:relative_trial_c +
    (1 | ind_id),
  family = binomial("cloglog"),
  REML = FALSE,
  data = msm_photeros,
  control = glmmTMBControl(optArgs = list(method = "L-BFGS-B"))
)

check_model(hazard_model_4)
check_collinearity(hazard_model_4)
summary(hazard_model_4)
drop1(hazard_model_4, test = "Chisq")

## photeros experience should be dropped if included in the model according to the Chi-sq test

hazard_model_5 <- glmmTMB(
  eaten ~ type + treatment_2 + relative_trial_c + photeros_attack_experience + time_point +
    type:photeros_attack_experience +
    treatment_2:relative_trial_c +
    (1 | ind_id),
  family = binomial("cloglog"),
  REML = FALSE,
  data = msm_photeros,
  control = glmmTMBControl(optArgs = list(method = "L-BFGS-B"))
)

compare_performance(hazard_model_4, hazard_model_5)
test_performance(hazard_model_4, hazard_model_5)

## in a model with either photeros attack OR luminescence experience
```

The model with the photeros_attack_experience is marginally better with the BF, but AICc are worse and the R\^2 are also worse. This supports that the model with the luminescence_experience is a better predictor of whether food is eaten than simply just attacking / experiencing the food. There does seem to be genus level differences, and we know that from looking at the prey response rates to the two different remaining genera, but the models are equivocal with and without it. Including it does not change the results.

### Looking at the Attack Model now

```{r Attack Model Data Prep}


msm_data_attack <- data_reduced %>%
  group_by(exp_id) %>%
  arrange(trial, order, time_point) %>%
  mutate(
    roll_num_attack = cumsum(num_attacks),
    behavior = case_when(
      eaten == "yes" ~ "Eaten",
      eaten == "no" &
        attacked == "yes" ~ "Rejected",
      eaten == "no" &
        attacked == "no" & roll_num_attack == 0 ~ "Unseen",
      eaten == "no" &
        attacked == "no" & roll_num_attack > 0 ~ "Avoided"
    ),
    behavior = as.factor(behavior),
    behavior = relevel(behavior, ref = "Eaten"),
    behavior_2 = case_when(
      eaten == "yes" ~ "Eaten",
      eaten == "no" &
        attacked == "yes" ~ "Rejected",
      eaten == "no" &
        attacked == "no" & roll_num_attack == 0 ~ "Ignored/Unseen"
    ),
    behavior_2 = as.factor(behavior_2),
    behavior_2 = relevel(behavior_2, ref = "Eaten"),
    behavior_1.3 = case_when(any(attacked == "yes", na.rm = TRUE) ~ "has attack", .default = "no attack"),
    treatment_2 = case_when(
      treatment == "frozen" ~ "dead",
      treatment == "boiled" ~ "dead",
      .default = treatment
    ),
    treatment_2 = as.factor(treatment_2),
    treatment_2 = relevel(treatment_2, ref = "live"),
    type = relevel(type, ref = "luminescent")
  ) %>% ungroup() %>% filter(type != "control") %>% droplevels()

trial_mean <- mean(msm_data_attack$relative_trial, na.rm = TRUE)
lum_mean   <- mean(msm_data_attack$luminescence_experience, na.rm = TRUE)
order_mean <- mean(msm_data_attack$order, na.rm = TRUE)

msm_data_attack <- msm_data_attack %>%
  mutate(
    relative_trial_c = relative_trial - mean(relative_trial, na.rm = TRUE),
    luminescence_experience_c = luminescence_experience - mean(luminescence_experience, na.rm = TRUE),
    order_c = order - mean(order, na.rm = TRUE),
    order = factor(
      order,
      ordered = TRUE,
      levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
    )
  )

```


```{r Attack Model}

attack_model_1 <- glmmTMB(
  attacked ~ type + #drop1
    treatment_2 + relative_trial_c + luminescence_experience_c +
    order + #drop1
    time_point +
    genus +
    #type:time_point + #too high of VIF
    #treatment_2:time_point + #too high of VIF
    order:relative_trial_c + #drop1
    type:luminescence_experience_c + #drop1
    type:relative_trial_c + #drop1
    treatment_2:relative_trial_c +
    treatment_2:luminescence_experience_c +
    (1 | ind_id / trial / sample_id),
  family = binomial("cloglog"),
  REML = FALSE,
  data = msm_data_attack %>% filter(genus != "Stegastes"),
  #using the entire data set with and without attacks
)

check_collinearity(attack_model_1)
drop1(attack_model_1, test = "Chisq")

# get rid of the following in round 1:
# - too high VIF: treatment_2:time_point, type:time_point

## keep the rest

```


```{r Attack Model Results}

summary(attack_model_1)

plot(effects::allEffects(attack_model_1), ylab = "Instantaneous risk of attack")

plot(effects::Effect(focal.predictors = "time_point", mod = attack_model_1),
     ylab = "Instantaneous risk of attack")
plot(effects::Effect(
  focal.predictors = c("treatment_2", "luminescence_experience_c"),
  mod = attack_model_1
),
ylab = "Instantaneous risk of attack")
plot(effects::Effect(
  focal.predictors = c("treatment_2", "relative_trial_c"),
  mod = attack_model_1
),
ylab = "Instantaneous risk of attack")

car::Anova(attack_model_1, "II")
check_model(attack_model_1)
check_collinearity(attack_model_1)
check_overdispersion(attack_model_1)
sjPlot::plot_model(
  attack_model_1,
  sort.est = "sort.all",
  pred.type = "fe",
  show.values = TRUE,
  value.offset = -0.35
)

```


```{r Attack Model Table}

sjPlot::tab_model(
  attack_model_1,
  show.intercept = FALSE,
  show.est = TRUE,
  show.dev = TRUE,
  show.loglik = TRUE,
  show.reflvl = TRUE
)

car::Anova(attack_model_1)

report::report(attack_model_1)

```


```{r Attack model: time point}


pred_grid <- expand.grid(
  type = levels(msm_data_attack$type),
  treatment_2 = levels(msm_data$treatment_2),
  relative_trial_c = seq(1:4),
  luminescence_experience_c = 0,
  #seq(0,max(msm_data_filtered$luminescence_experience)),
  time_point = levels(msm_data_attack$time_point),
  order =  levels(msm_data_attack$order),
  genus = c("Abudefduf", "Phaeoptyx")
)

pred_grid$time_point <- factor(
  pred_grid$time_point,
  levels = levels(msm_data_attack$time_point),
  ordered = TRUE
)

X <- model.matrix(lme4::nobars(formula(attack_model_1, component = "cond"))[-2], pred_grid)

beta_hat <- fixef(attack_model_1)$cond
V_beta <- vcov(attack_model_1)$cond

set.seed(123)

nsim <- 1000

beta_sim <- MASS::mvrnorm(nsim, mu = beta_hat, Sigma = V_beta)

eta_sim <- X %*% t(beta_sim)

haz_sim <- 1 - exp(-exp(eta_sim))

sim_df <- bind_cols(pred_grid, as.data.frame(haz_sim))

names(sim_df)[(ncol(pred_grid) + 1):ncol(sim_df)] <-
  paste0("sim_", seq_len(nsim))

plot_df <- sim_df %>%
  arrange(genus,
          order,
          treatment_2,
          relative_trial_c,
          luminescence_experience_c,
          time_point) %>%
  group_by(genus,
           order,
           treatment_2,
           relative_trial_c,
           luminescence_experience_c) %>%
  mutate(across(starts_with("sim_"), ~ 1 - cumprod(1 - .x))) %>%
  ungroup()

plot_df <- plot_df %>%
  rowwise() %>%
  mutate(
    prob = mean(c_across(starts_with("sim_"))),
    lower = quantile(c_across(starts_with("sim_")), 0.025),
    upper = quantile(c_across(starts_with("sim_")), 0.975)
  ) %>%
  ungroup() %>% mutate(type = case_when(type == "luminescent" ~ "bioluminescent",.default = type))

plot_df <- plot_df %>%
  mutate(condition = interaction(str_to_title(type), str_to_title(treatment_2)))

FigureS6_1 <- ggplot(plot_df, aes(x = time_point, y = prob)) +
  geom_ribbon(aes(
    ymin = lower,
    ymax = upper,
    group = condition,
    fill = condition
  ),
  alpha = 0.1) +
  geom_line(linewidth = 1.2, aes(group = condition, color = condition)) +
  geom_point(
    size = 3,
    color = "black",
    shape = 21,
    aes(fill = condition, group = condition)
  ) +
  coord_cartesian(ylim = c(0, 1)) +
  labs(
    title = "Cumulative probability of being attacked over time, across presentation order, and across trials",
    subtitle = "Separated by predator genus, prey treatment, and prey type",
    x = "Time point",
    y = "Cumulative probability of being attacked",
  ) +
  theme_bw(base_size = 20) +
  facet_nested(
    #genus + relative_trial_c ~ order,
    genus + treatment_2 + type ~ relative_trial_c + order,
    nest_line = element_line(),
    strip =
      strip_nested(
        by_layer_y = TRUE,
        text_y = list(element_text(face = "italic"), element_text(face = "plain"))
      )
  ) +
  theme(strip.background = element_blank()) +
  scale_color_manual(
    name = "Prey Type",
    breaks = c(
      'Bioluminescent.Live','Nonluminescent.Live',
      'Bioluminescent.Anaesthetized','Nonluminescent.Anaesthetized',
      'Bioluminescent.Dead','Nonluminescent.Dead'
    ),
    values = c(
      'Bioluminescent.Live' = "#376795",
      "Nonluminescent.Live" = "#e76254",
      'Bioluminescent.Anaesthetized' = "#528fad",
      'Nonluminescent.Anaesthetized' = "#ef7a57",
      'Bioluminescent.Dead' = "#72bcd5",
      'Nonluminescent.Dead' = "#f7aa58"
    )
  ) +
  
  scale_fill_manual(
    name = "Prey Type",
    breaks = c(
      'Bioluminescent.Live','Nonluminescent.Live',
      'Bioluminescent.Anaesthetized','Nonluminescent.Anaesthetized',
      'Bioluminescent.Dead','Nonluminescent.Dead'
    ),
    values = c(
      'Bioluminescent.Live' = "#376795",
      "Nonluminescent.Live" = "#e76254",
      'Bioluminescent.Anaesthetized' = "#528fad",
      'Nonluminescent.Anaesthetized' = "#ef7a57",
      'Bioluminescent.Dead' = "#72bcd5",
      'Nonluminescent.Dead' = "#f7aa58"
    )
  )

ggsave(
  plot = FigureS6_1,
  filename = "./submissions/plots/attack_genus_order_type_treatment.jpeg",
  dpi = 300,
  units = "px",
  height = 7000,
  width = 13000,
  limitsize = TRUE,
  device = "jpeg"
)
```

Plotting the cumulative probability of being eaten for the effects of the attack model as well. I need the following plots:

1)  Type x Treatment
2)  Type x Experience x Trial
3)  Treatment x Experience x Trial


```{r Attack model proportion graph}
## ----------------------------
## 1. model-based survival predictions
## ----------------------------

pred_grid <- expand.grid(
  type = levels(msm_data_attack$type),
  treatment_2 = levels(msm_data$treatment_2),
  relative_trial_c = 0,
  #seq(1:4),
  luminescence_experience_c = 0,
  #seq(0,max(msm_data_filtered$luminescence_experience)),
  time_point = levels(msm_data_attack$time_point),
  order =  levels(msm_data_attack$order),
  genus = c("Abudefduf", "Phaeoptyx")
)

## model matrix for fixed effects
X <- model.matrix(delete.response(terms(attack_model_1)), pred_grid)

## fixed-effect estimates and covariance matrix
beta_hat <- fixef(attack_model_1)$cond
V_beta <- vcov(attack_model_1)$cond

## simulate from fixed-effect uncertainty
set.seed(123)

nsim <- 1000

beta_sim <- MASS::mvrnorm(n = nsim, mu = beta_hat, Sigma = V_beta)

## convert linear predictor -> cloglog hazard
eta_sim <- X %*% t(beta_sim)

haz_sim <- 1 - exp(-exp(eta_sim))

## bind simulations to pred_grid
sim_df <- bind_cols(pred_grid, as.data.frame(haz_sim))

names(sim_df)[(ncol(pred_grid) + 1):ncol(sim_df)] <- paste0("sim_", seq_len(nsim))

## ----------------------------
## 2. hazards -> cumulative probability eaten within trial
## ----------------------------

plot_df <- sim_df %>%
  arrange(genus, order, type, treatment_2, relative_trial_c, time_point) %>%
  group_by(genus, order, type, treatment_2, relative_trial_c) %>%
  mutate(across(starts_with("sim_"), ~ 1 - cumprod(1 - .x))) %>%
  filter(time_point == levels(time_point)[nlevels(time_point)]) %>%
  ungroup() %>%
  group_by(type, treatment_2) %>%
  summarise(across(starts_with("sim_"), mean), .groups = "drop") %>%
  rowwise() %>%
  mutate(
    pred  = mean(c_across(starts_with("sim_"))),
    lower = quantile(c_across(starts_with("sim_")), 0.025),
    upper = quantile(c_across(starts_with("sim_")), 0.975)
  ) %>%
  ungroup() %>%
  mutate(
    type = str_to_title(type),
    type = case_when(type == "Luminescent" ~ "Bioluminescent",.default = type),
    treatment_2 = str_to_title(treatment_2),
    type = fct_relevel(type, c("Bioluminescent", "Nonluminescent")),
    treatment_2 = fct_relevel(treatment_2, c("Live", "Anaesthetized", "Dead")),
    condition = interaction(type, treatment_2)
  )

## ----------------------------
## 3. pairwise contrasts on final cumulative probability eaten
## ----------------------------

sim_long <- plot_df %>%
  dplyr::select(type, treatment_2, condition, starts_with("sim_")) %>%
  tidyr::pivot_longer(cols = starts_with("sim_"),
                      names_to = "sim",
                      values_to = "pred_eaten")

contrast_df <- sim_long %>%
  dplyr::select(condition, sim, pred_eaten) %>%
  tidyr::pivot_wider(names_from = condition, values_from = pred_eaten)

pairwise_contrasts <- combn(levels(plot_df$condition), 2, simplify = FALSE) %>%
  purrr::map_dfr(function(x) {
    diff <- contrast_df[[x[1]]] - contrast_df[[x[2]]]
    
    tibble::tibble(
      group1 = x[1],
      group2 = x[2],
      estimate = mean(diff),
      lower = quantile(diff, 0.025),
      upper = quantile(diff, 0.975),
      significant = lower > 0 | upper < 0
    )
  })

library(multcompView)

p_mat <- matrix(
  1,
  nrow = nlevels(plot_df$condition),
  ncol = nlevels(plot_df$condition),
  dimnames = list(levels(plot_df$condition), levels(plot_df$condition))
)

for (i in seq_len(nrow(pairwise_contrasts))) {
  p_mat[pairwise_contrasts$group1[i], pairwise_contrasts$group2[i]] <-
    ifelse(pairwise_contrasts$significant[i], 0.01, 1)
  
  p_mat[pairwise_contrasts$group2[i], pairwise_contrasts$group1[i]] <-
    ifelse(pairwise_contrasts$significant[i], 0.01, 1)
}

letters_df <- multcompView::multcompLetters(p_mat, threshold = 0.05)$Letters %>%
  tibble::enframe(name = "condition", value = ".group")

plot_df <- plot_df %>%
  left_join(letters_df, by = "condition")

## visualizing the distributions of the simulations for each group
sim_long %>% ggplot(aes(x = pred_eaten)) +
  geom_histogram(aes(color = condition, fill = condition),
                 alpha = 0.15,
                 position = "identity") +
  theme_bw() + xlab("Cumulative Probability of Being Attacked (n = 1000)") + ylab("Count")


## ----------------------------
## 2. raw observed proportions for bars
## ----------------------------

bars_df <- msm_data_attack %>% filter(genus != "Stegastes") %>%
  group_by(exp_id) %>%
  arrange(ind_id, trial, order, time_point, .by_group = TRUE) %>%
  filter(time_point == last(time_point)) %>%
  ungroup() %>%
  mutate(
    sample_id = case_when(
      str_detect(sample_id, "photeros.boiled") ~ "photeros.dead",
      str_detect(sample_id, "photeros.frozen") ~ "photeros.dead",
      str_detect(sample_id, "skogsbergia.boiled") ~ "skogsbergia.dead",
      str_detect(sample_id, "skogsbergia.frozen") ~ "skogsbergia.dead",
      TRUE ~ sample_id
    )
  ) %>%
  group_by(sample_id) %>%
  summarise(
    prop_attacked = mean(behavior_1.3 == "has attack", na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(
    type = case_when(
      str_detect(sample_id, "photeros") ~ "Bioluminescent",
      str_detect(sample_id, "skogsbergia") ~ "Nonluminescent",
      TRUE ~ NA_character_
    ),
    treatment_2 = case_when(
      str_detect(sample_id, "live") ~ "Live",
      str_detect(sample_id, "anaesthetized") ~ "Anaesthetized",
      str_detect(sample_id, "dead") ~ "Dead",
      TRUE ~ NA_character_
    ),
    type = fct_relevel(type, c("Bioluminescent", "Nonluminescent")),
    treatment_2 = fct_relevel(treatment_2, c("Live", "Anaesthetized", "Dead")),
    condition = interaction(type, treatment_2)
  ) %>%
  filter(!is.na(type), !is.na(treatment_2))

## ----------------------------
## 3. plot
## ----------------------------

pd <- position_dodge2(width = 0.75)

FigureS666 <- ggplot() +
  geom_col(
    data = bars_df,
    aes(
      x = treatment_2,
      y = prop_attacked,
      fill = condition,
      colour = condition
    ),
    position = pd,
    alpha = 1,
    width = 0.75,
    linewidth = 0.90
  ) +
  geom_text(
    data = bars_df,
    aes(
      x = treatment_2,
      y = -0.05,
      label = paste0("n = ", n),
      colour = condition
    ),
    position = pd,
    size = 5
  ) +
  geom_linerange(
    data = plot_df,
    aes(x = treatment_2, ymin = lower, ymax = upper),
    color = "black",
    position = pd,
    linewidth = 1.2
  ) +
  geom_point(
    data = plot_df,
    aes(
      x = treatment_2,
      y = pred,
      fill = condition,
      shape = condition
    ),
    color = "black",
    position = pd,
    size = 6.5,
    stroke = 1.2
  ) +
  geom_text(
    data = plot_df,
    aes(x = treatment_2, y = 1.05, label = .group),
    position = pd,
    size = 10,
    fontface = "bold"
  ) +
  scale_shape_manual(
    name = "Prey Type",
    breaks = c(
      'Bioluminescent.Live', 'Nonluminescent.Live', 'Bioluminescent.Anaesthetized', 'Nonluminescent.Anaesthetized', 'Bioluminescent.Dead', 'Nonluminescent.Dead'
    ),
    values = c(
      'Bioluminescent.Live' = 25,
      "Nonluminescent.Live" = 24,
      'Bioluminescent.Anaesthetized' = 25,
      'Nonluminescent.Anaesthetized' = 24,
      'Bioluminescent.Dead' = 25,
      'Nonluminescent.Dead' = 24
    )
  ) +
  scale_color_manual(
    name = "Prey Type",
    breaks = c(
      'Bioluminescent.Live', 'Nonluminescent.Live', 'Bioluminescent.Anaesthetized', 'Nonluminescent.Anaesthetized', 'Bioluminescent.Dead', 'Nonluminescent.Dead'
    ),
    values = c(
      'Bioluminescent.Live' = "#376795",
      "Nonluminescent.Live" = "#e76254",
      'Bioluminescent.Anaesthetized' = "#528fad",
      'Nonluminescent.Anaesthetized' = "#ef7a57",
      'Bioluminescent.Dead' = "#72bcd5",
      'Nonluminescent.Dead' = "#f7aa58"
    )
  ) +
  scale_fill_manual(
    name = "Prey Type",
    breaks = c(
      'Bioluminescent.Live', 'Nonluminescent.Live', 'Bioluminescent.Anaesthetized', 'Nonluminescent.Anaesthetized', 'Bioluminescent.Dead', 'Nonluminescent.Dead'
    ),
    values = c(
      'Bioluminescent.Live' = "#376795",
      "Nonluminescent.Live" = "#e76254",
      'Bioluminescent.Anaesthetized' = "#528fad",
      'Nonluminescent.Anaesthetized' = "#ef7a57",
      'Bioluminescent.Dead' = "#72bcd5",
      'Nonluminescent.Dead' = "#f7aa58"
    )
  ) +
  coord_cartesian(ylim = c(-0.07, 1.1)) +
  scale_y_continuous(breaks = c(0.25, 0.50, 0.75, 1)) +
  labs(title = "Proportion of trials where predators attacked prey \u2265 1", x = "Treatment", y = "Proportion attacked (bars)\nCumulative predicted probability of being attacked + 95% CI (triangles + whiskers)") +
  theme_bw(base_size = 20)

FigureS666

```


```{r Attack model: Type x Trial x Experience}

pred_grid <- expand.grid(
  type = levels(msm_data_attack$type),
  treatment_2 = levels(msm_data_attack$treatment_2),
  relative_trial_c = seq(1, 4),
  luminescence_experience_c = seq(0, 18),
  time_point = levels(msm_data_attack$time_point),
  order = levels(msm_data_attack$order),
  genus = c("Abudefduf", "Phaeoptyx")
)

pred_grid$time_point <- factor(
  pred_grid$time_point,
  levels = levels(msm_data_attack$time_point),
  ordered = TRUE
)

X <- model.matrix(lme4::nobars(formula(attack_model_1, component = "cond"))[-2], pred_grid)

beta_hat <- fixef(attack_model_1)$cond
V_beta <- vcov(attack_model_1)$cond

set.seed(123)

nsim <- 1000

beta_sim <- MASS::mvrnorm(nsim, mu = beta_hat, Sigma = V_beta)

eta_sim <- X %*% t(beta_sim)

haz_sim <- 1 - exp(-exp(eta_sim))

sim_df <- bind_cols(pred_grid, as.data.frame(haz_sim))

names(sim_df)[(ncol(pred_grid) + 1):ncol(sim_df)] <-
  paste0("sim_", seq_len(nsim))

plot_df <- sim_df %>%
  arrange(
    genus,
    type,
    order,
    treatment_2,
    relative_trial_c,
    luminescence_experience_c,
    time_point
  ) %>%
  group_by(genus,
           type,
           order,
           treatment_2,
           relative_trial_c,
           luminescence_experience_c,
  ) %>%
  mutate(across(starts_with("sim_"), ~ 1 - cumprod(1 - .x))) %>%
  filter(time_point == levels(time_point)[nlevels(time_point)]) %>%
  ungroup() %>%
  group_by(type, luminescence_experience_c, relative_trial_c) %>%
  summarise(across(starts_with("sim_"), mean), .groups = "drop") %>%
  rowwise() %>%
  mutate(
    prob  = mean(c_across(starts_with("sim_"))),
    lower = quantile(c_across(starts_with("sim_")), 0.025),
    upper = quantile(c_across(starts_with("sim_")), 0.975)
  ) %>%
  ungroup() %>%
  mutate(type = str_to_title(type), type = case_when(type == "Luminescent" ~ "Bioluminescent",.default = type),)

FigureS6_2 <- ggplot(plot_df,
                     aes(
                       x = as.numeric(luminescence_experience_c),
                       y = prob,
                       colour = type,
                       fill = type,
                       group = type
                     )) +
  geom_ribbon(aes(ymin = lower, ymax = upper),
              alpha = 0.15,
              colour = NA) +
  geom_line(linewidth = 1.2, aes(color = type)) +
  coord_cartesian(ylim = c(0, 1)) +
  labs(
    title = "Cumulative probability of being attacked over bioluminescence experience and across trials",
    subtitle = "Separated by prey type",
    x = "Biouminescence experience",
    y = "Cumulative probability of being attacked",
  ) +
  theme_bw(base_size = 20) + facet_grid(. ~ as.numeric(relative_trial_c)) +
  theme(strip.background = element_blank()) +
  scale_color_manual(
    name = "Prey Type",
    values = c(
      "Bioluminescent" = "#376795",
      "Nonluminescent" = "#e76254"
    )
  ) +
  scale_fill_manual(
    name = "Prey Type",
    values = c(
      "Bioluminescent" = "#376795",
      "Nonluminescent" = "#e76254"
    )
  )

FigureS6_2
```


```{r Attack model: Treatment x Trial x Experience}

pred_grid <- expand.grid(
  type = levels(msm_data_attack$type),
  treatment_2 = levels(msm_data_attack$treatment_2),
  relative_trial_c = seq(1, 4),
  luminescence_experience_c = seq(0, 18),
  time_point = levels(msm_data_attack$time_point),
  order = levels(msm_data_attack$order),
  genus = c("Abudefduf", "Phaeoptyx")
)

pred_grid$time_point <- factor(
  pred_grid$time_point,
  levels = levels(msm_data_attack$time_point),
  ordered = TRUE
)

X <- model.matrix(lme4::nobars(formula(attack_model_1, component = "cond"))[-2], pred_grid)

beta_hat <- fixef(attack_model_1)$cond
V_beta <- vcov(attack_model_1)$cond

set.seed(123)

nsim <- 1000

beta_sim <- MASS::mvrnorm(nsim, mu = beta_hat, Sigma = V_beta)

eta_sim <- X %*% t(beta_sim)

haz_sim <- 1 - exp(-exp(eta_sim))

sim_df <- bind_cols(pred_grid, as.data.frame(haz_sim))

names(sim_df)[(ncol(pred_grid) + 1):ncol(sim_df)] <-
  paste0("sim_", seq_len(nsim))

plot_df <- sim_df %>%
  arrange(
    genus,
    type,
    order,
    treatment_2,
    relative_trial_c,
    luminescence_experience_c,
    time_point
  ) %>%
  group_by(genus,
           type,
           order,
           treatment_2,
           relative_trial_c,
           luminescence_experience_c,
  ) %>%
  mutate(across(starts_with("sim_"), ~ 1 - cumprod(1 - .x))) %>%
  filter(time_point == levels(time_point)[nlevels(time_point)]) %>%
  ungroup() %>%
  group_by(treatment_2, luminescence_experience_c, relative_trial_c) %>%
  summarise(across(starts_with("sim_"), mean), .groups = "drop") %>%
  rowwise() %>%
  mutate(
    prob  = mean(c_across(starts_with("sim_"))),
    lower = quantile(c_across(starts_with("sim_")), 0.025),
    upper = quantile(c_across(starts_with("sim_")), 0.975)
  ) %>%
  ungroup() %>%
  mutate(type = str_to_title(treatment_2))

FigureS6_1.3 <- plot_df %>% mutate(
  treatment_2 = case_when(
    treatment_2 == "live" ~ "Live",
    treatment_2 == "anaesthetized" ~ "Anaesthetized",
    treatment_2 == "dead" ~ "Dead"
  ),
  treatment_2 = factor(treatment_2, labels = c("Live", "Anaesthetized", "Dead")),
) %>% ggplot(
  aes(
    x = as.numeric(luminescence_experience_c),
    y = prob,
    colour = treatment_2,
    fill = treatment_2,
    group = treatment_2
  )
) +
  geom_ribbon(aes(ymin = lower, ymax = upper),
              alpha = 0.15,
              colour = NA) +
  geom_line(linewidth = 1.2, aes(color = treatment_2)) +
  coord_cartesian(ylim = c(0, 1)) +
  labs(
    title = "Cumulative probability of being attacked over bioluminescence experience and across trials",
    subtitle = "Separated by prey treatment",
    x = "Bioluminescence experience",
    y = "Cumulative probability of being attacked"
  ) +
  theme_bw(base_size = 20) + facet_grid(treatment_2 ~ as.numeric(relative_trial_c)) +
  theme(strip.background = element_blank()) +
  scale_color_manual(
    name = "Prey Treatment",
    values = c(
      "Live" = "#302F8AFF",
      "Anaesthetized" = "#602F8AFF",
      "Dead" = "#9B4F82FF"
    )
  ) +
  scale_fill_manual(
    name = "Prey Treatment",
    values = c(
      "Live" = "#302F8AFF",
      "Anaesthetized" = "#602F8AFF",
      "Dead" = "#9B4F82FF"
    )
  )

FigureS6_1.3
```


```{r Saving them as a plot}

FigS6 <- ggarrange(
  FigureS666,
  FigureS6_2,
  FigureS6_1.3,
  nrow = 3,
  labels = c("A", "B", "C"),
  font.label = list(size = 22),
  heights = c(2.65, 1.5, 2)
)

ggsave(
  filename = "FigureS6.jpeg",
  plot = FigS6,
  device = "jpeg",
  path = "./submissions/plots/",
  width = 15,
  height = 24,
  units = "in",
  dpi = 300,
  limitsize = TRUE
)
```

